### 2020-06-29 paper.tex
initial commit

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{multicol}

\def\code#1{\texttt{#1}}

\usepackage{biblatex}
\addbibresource{sources.bib}

\evensidemargin=0in
\oddsidemargin=0in
\textwidth=6.5in
\textheight=9.0in

\title{Sorting by Reversals}
\author{Jonas Hübotter}

\begin{document}

\maketitle

\printbibliography

\end{document}

### 2020-06-29 slides.tex
initial commit

\documentclass{beamer}
%
% Choose how your presentation looks.
%
% For more themes, color themes and font themes, see:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
%
\mode<presentation>
{
  \usetheme{default}      % or try Darmstadt, Madrid, Warsaw, ...
  \usecolortheme{default} % or try albatross, beaver, crane, ...
  \usefonttheme{default}  % or try serif, structurebold, ...
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
  \setbeamertemplate{footline}[frame number]
  \setbeamertemplate{itemize items}[circle]
  \setbeamertemplate{theorems}[numbered]
  \setbeamercolor*{structure}{bg=white,fg=blue}
  \setbeamerfont{block title}{size=\normalsize}
}

\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{idea}[theorem]{Idea}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{aligned-overset}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{multicol}
\usepackage{stmaryrd}
\usepackage{tabularx}

\renewcommand\tabularxcolumn[1]{m{#1}}
\newcolumntype{R}{>{\raggedleft\arraybackslash}X}

\usepackage{tikz}
\tikzstyle{node}=[fill=white, draw=black, shape=circle, minimum size=0.75cm]
\tikzstyle{blue node}=[fill=white, draw=blue, shape=circle]
\tikzstyle{red node}=[fill=white, draw=red, shape=circle]
\tikzstyle{blue edge}=[-, draw=blue]
\tikzstyle{red edge}=[-, draw=red]
\pgfdeclarelayer{nodelayer}
\pgfdeclarelayer{edgelayer}
\pgfsetlayers{nodelayer,edgelayer}

\def\code#1{\texttt{\frenchspacing#1}}
\def\padding{\vspace{0.5cm}}
\def\spadding{\vspace{0.25cm}}
\def\b{\textcolor{blue}}
\def\r{\textcolor{red}}

% fix for \pause in align
\makeatletter
\let\save@measuring@true\measuring@true
\def\measuring@true{%
  \save@measuring@true
  \def\beamer@sortzero##1{\beamer@ifnextcharospec{\beamer@sortzeroread{##1}}{}}%
  \def\beamer@sortzeroread##1<##2>{}%
  \def\beamer@finalnospec{}%
}
\makeatother

\usepackage{biblatex}
\addbibresource{sources.bib}

\title[Sorting by Reversals]{Sorting by Reversals}
\author{Jonas Hübotter}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

% Uncomment these lines for an automatically generated outline.
\begin{frame}{Outline}
 \tableofcontents
\end{frame}

\section{Introduction}

\subsection{Symmetric group}

\begin{frame}

\begin{definition}
We define the \b{symmetric group} $\langle S_n, \circ \rangle$ as the group whose elements are all bijections over $[1,n]$ \pause with
\begin{align*}
    S_n = \{(0\ \pi_1\ \dots\ \pi_n\ n+1) \mid \{\pi_1, \dots, \pi_n\} = [1,n]\}
\end{align*}
where $\pi_i = \pi(i)$, $\pi_0 = 0$, and $\pi_{n+1} = n+1$.\pause\spadding

$\pi \in S_n$ is a \b{permutation}.\pause\spadding

$id = (0\ 1\ \dots\ n\ n+1) \in S_n$ is the \textit{identity permutation}.
\end{definition}

\end{frame}

\begin{frame}

\begin{definition}
A \b{reversal} $\rho(i,j) \in S_n$ is defined as
\[
    \rho(i,j) = (0\ 1\ \cdots\ i-1\ \r j\ \r{j-1}\ \r \cdots\ \r{i+1}\ \r i\ j+1\ \cdots\ n\ n+1)
\]
for some $j \geq i$.
\end{definition}\pause

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. \par
Then
\begin{align*}
    \pi \circ \rho(1,4) = (0\ \r{3\ 1\ 2\ 4}\ 5).
\end{align*}
\end{exampleblock}

\end{frame}

\subsection{Reversal distance problem}

\begin{frame}

\begin{definition}[reversal distance problem]
Given two permutations $\sigma, \tau \in S_n$ find a sequence of reversals $\rho_1, \dots, \rho_d \in S_n$ such that
\begin{align*}
    \sigma \circ \rho_1 \circ \cdots \circ \rho_d = \tau
\end{align*}
and $d$ is minimal. \pause\par\padding
$d$ is called \b{reversal distance} between $\sigma$ and $\tau$. \cite{Kececioglu1995}
\end{definition}

\end{frame}

\section{MIN-SBR}

\begin{frame}

Observation: The reversal distance between $\sigma$ and $\tau$ is the same as the reversal distance between $\tau^{-1} \circ \sigma$ and $id$. \pause\padding

\begin{definition}[MIN-SBR]
Let $\pi = \tau^{-1} \circ \sigma \in S_n$. \par\pause
\textit{Sorting by Reversals} is the problem of finding a sequence of reversals $\rho_1, \dots, \rho_d \in S_n$ such that
\begin{align*}
    \pi \circ \rho_1 \circ \cdots \circ \rho_d = id
\end{align*}
and $d$ is minimal. \cite{Kececioglu1995}
\end{definition}

\end{frame}

\begin{frame}

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$.\pause
\begin{align*}
    \pi \circ \rho(1,3) &= (0\ \r{1\ 2\ 4}\ 3\ 5) \\ \pause
    \pi \circ \rho(1,3) \circ \rho(3,4) &= (0\ 1\ 2\ \r{3\ 4}\ 5) = id
\end{align*}\pause
$\implies d(\pi) \leq 2$.
\end{exampleblock}\pause\padding

\begin{algorithm}[Watterson]
For each element at position $i$, perform $\rho(i, \pi_i^{-1})$ if $i \neq \pi_i$. \cite{Kececioglu1995}
\end{algorithm}\pause\padding

\begin{corollary}[tight upper bound, \citeauthor*{Kececioglu1995}]
$d(\pi) \leq n - 1$ for all $\pi \in S_n$.
\end{corollary}

\end{frame}

\section{Breakpoint graph}

\begin{frame}

\begin{definition}
Let $i \sim j$ if $|i - j| = 1$. \par\pause
A pair of consecutive elements $\pi_i$ and $\pi_j$ is
\begin{itemize}
    \item an \b{adjacency} if $\pi_i \sim \pi_j$; and
    \item a \b{breakpoint} if $\pi_i \not\sim \pi_j$.
\end{itemize} \par\pause\spadding
$b(\pi)$ denotes the number of breakpoints in $\pi \in S_n$.
\end{definition}\pause\padding

Observation: $b(\pi) = 0$ iff $\pi = id$ and any reversal can at most eliminate two breakpoints. \pause\spadding

\begin{corollary}[lower bound, \citeauthor*{Kececioglu1995}]
$d(\pi) \geq \left\lceil \frac{b(\pi)}{2} \right\rceil$ for all $\pi \in S_n$.
\end{corollary}

\end{frame}

\begin{frame}

\begin{definition}[breakpoint graph, \citeauthor*{Bafna1996}]
Let $G(\pi) = (V, E)$ with
\begin{itemize}
    \item vertices $V = [0, n+1]$ representing the elements of $\pi$\pause; and
    \item edges $E = \r R \cup \b B$ with
    \begin{itemize}
        \item a \r{red} edge for every breakpoint in $\pi$; and
        \item a \b{blue} edge for every adjacency in $\pi$.
    \end{itemize}
\end{itemize}
\end{definition}

\end{frame}

\begin{frame}

\begin{exampleblock}{Example}
Let $\pi = (0\ \r\mid\ 4\ \r\mid\ 2\ 1\ \r\mid\ 3\ \r\mid\ 5) \in S_4$.

\begin{center}
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\end{exampleblock}\pause

Observation: Each vertex has an equal number of incident red and blue edges. \pause\spadding

\begin{corollary}[\citeauthor*{Bafna1996}]
$G(\pi)$ can be decomposed into edge-disjoint alternating cycles.
\end{corollary}

\end{frame}

\begin{frame}

\begin{definition}
A reversal is called \b{$k$-reversal} if it removes $k$ breakpoints. \pause\spadding

A reversal \b{acts on} two red edges of $G\pi)$ if those two edges represent the breakpoints that are split apart by the reversal. \pause\spadding

An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges.
\end{definition}

\end{frame}

\begin{frame}

\begin{definition}
We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$. \pause\spadding

An alternating cycle in $G(\pi)$ is \b{directed} with respect to $\pi$ if it is possible to walk along the whole cycle traversing each red edge in the direction of its orientation.
\end{definition}

\end{frame}

\begin{frame}

\begin{proposition}[\citeauthor*{Caprara1997}]
An alternating cycle in $G(\pi)$ is oriented iff it is undirected.
\end{proposition}\pause

\begin{exampleblock}{Example}
\begin{center}
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (6) at (-3.75, 0) {0};
		\node [style=node] (7) at (-2.25, 0) {3};
		\node [style=node] (8) at (-0.75, 0) {4};
		\node [style=node] (9) at (0.75, 0) {1};
		\node [style=node] (10) at (2.25, 0) {2};
		\node [style=node] (11) at (3.75, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (6) to (7);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (10) to (11);
		\draw [style=blue edge, bend right] (6) to (9);
		\draw [style=blue edge, bend left] (7) to (10);
		\draw [style=blue edge, bend right] (8) to (11);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}\pause
$\implies$ directed and unoriented.
\end{exampleblock}

\end{frame}

\begin{frame}

\begin{lemma}[\citeauthor*{Christie1998}]
\label{lem:2}
An oriented alternating cycle is a $2$-cycle iff the corresponding reversal acting on both red edges is a $2$-reversal.
\end{lemma}\pause

\begin{proof}
Let $i \sim i'$ and $j \sim j'$.

\begin{center}
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-1.25, 0) {i};
		\node [style=node] (1) at (1.25, 0) {j'};
		\node [style=node] (2) at (3, 0) {i'};
		\node [style=node] (3) at (-3, 0) {j};
		\node (4) at (0, 0) {$\dots$};
		\node (5) at (-4.25, 0) {$\dots$};
		\node (6) at (4.25, 0) {$\dots$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=blue edge, bend right] (3) to (1);
		\draw [style=blue edge, bend left] (0) to (2);
		\draw [style=red edge] (3) to (0);
		\draw [style=red edge] (1) to (2);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\end{proof}

\end{frame}

\begin{frame}

Let $c(\pi)$ denote the maximum number of alternating cycles in any alternating cycle decomposition of $G(\pi)$. \pause\padding

\begin{theorem}[\citeauthor*{Bafna1996}]
\label{thm:1}
Let $\pi, \rho \in S_n$ and $\rho$ be a reversal. \pause Then
\begin{align*}
    b(\pi) - b(\pi \circ \rho) + c(\pi \circ \rho) - c(\pi) \leq 1.
\end{align*}
\end{theorem}

\end{frame}

\begin{frame}

\begin{proof}
We consider each case $b(\pi) - b(\pi \circ \rho) \in [-2, 2]$ separately.\pause
\begin{enumerate}
    \item A $2$-reversal removes at least one alternating cycle from the maximum alternating cycle decomposition.
        \begin{tabularx}{\textwidth}{XcX}
            \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
            	\begin{pgfonlayer}{nodelayer}
            		\node [style=node] (0) at (-1.25, 0) {i};
            		\node [style=node] (1) at (1.25, 0) {j'};
            		\node [style=node] (2) at (3, 0) {i'};
            		\node [style=node] (3) at (-3, 0) {j};
            		\node (4) at (0, 0) {$\dots$};
            	\end{pgfonlayer}
            	\begin{pgfonlayer}{edgelayer}
            		\draw [style=blue edge, bend right] (3) to (1);
            		\draw [style=blue edge, bend left] (0) to (2);
            		\draw [style=red edge] (3) to (0);
            		\draw [style=red edge] (1) to (2);
            	\end{pgfonlayer}
            \end{tikzpicture} & $\rightarrow$ &
            \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
            	\begin{pgfonlayer}{nodelayer}
            		\node [style=node] (0) at (-1.25, 0) {j'};
            		\node [style=node] (1) at (1.25, 0) {i};
            		\node [style=node] (2) at (3, 0) {i'};
            		\node [style=node] (3) at (-3, 0) {j};
            		\node (4) at (0, 0) {$\dots$};
            	\end{pgfonlayer}
            \end{tikzpicture}
        \end{tabularx}\pause

    \item A $1$-reversal does not add an alternating cycle to the maximum alternating cycle decomposition.
        \begin{tabularx}{\textwidth}{XcX}
            \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
            	\begin{pgfonlayer}{nodelayer}
            		\node [style=node] (0) at (-1.25, 0) {i};
            		\node [style=node] (1) at (1.25, 0) {j'};
            		\node [style=node] (2) at (3, 0) {k};
            		\node [style=node] (3) at (-3, 0) {j};
            		\node (4) at (0, 0) {$\dots$};
            	\end{pgfonlayer}
            	\begin{pgfonlayer}{edgelayer}
            		\draw [style=blue edge, bend right] (3) to (1);
            		\draw [style=red edge] (3) to (0);
            		\draw [style=red edge] (1) to (2);
            		\draw [dashed, bend left] (0) to (2);
            	\end{pgfonlayer}
            \end{tikzpicture} & $\rightarrow$ &
            \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
            	\begin{pgfonlayer}{nodelayer}
            		\node [style=node] (0) at (-1.25, 0) {j'};
            		\node [style=node] (1) at (1.25, 0) {i};
            		\node [style=node] (2) at (3, 0) {k};
            		\node [style=node] (3) at (-3, 0) {j};
            		\node (4) at (0, 0) {$\dots$};
            	\end{pgfonlayer}
            	\begin{pgfonlayer}{edgelayer}
            		\draw [style=red edge] (1) to (2);
            		\draw [dashed, bend left] (1) to (2);
            	\end{pgfonlayer}
            \end{tikzpicture}
        \end{tabularx}
\end{enumerate}\pause
Proof for other cases similar.
\end{proof}

\end{frame}

\begin{frame}

\begin{theorem}[lower bound, \citeauthor*{Bafna1996}]
\label{thm:2}
Let $\pi \in S_n$. Then
\begin{align*}
    d(\pi) \geq b(\pi) - c(\pi).
\end{align*}
\end{theorem}\pause

\begin{proof}\renewcommand{\qedsymbol}{}
Let $\pi_t = \pi, \pi_0 = id$ and $\rho_1, \dots, \rho_t$ a shortest sequence of reversals from $\pi_t$ to $\pi_0$. \par\pause
Then
\begin{align*}
    d(\pi_i) &= d(\pi_{i-1}) + 1 \\ \pause
             \overset{(\ref{thm:1})}&{\geq} d(\pi_{i-1}) + b(\pi_i) - b(\pi_{i-1}) + c(\pi_{i-1}) - c(\pi_i).
\end{align*}
\end{proof}

\end{frame}

\begin{frame}

\begin{proof}[Proof (cont.)]
Therefore
\begin{align*}
    d(\pi_i) - (b(\pi_i) - c(\pi_i)) &\geq d(\pi_{i-1}) - (b(\pi_{i-1}) - c(\pi_{i-1})) \\ \pause
                                     &\geq d(\pi_0) - (b(\pi_0) - c(\pi_0))\pause = 0
\end{align*}\pause
Let $i = t$. Then
\begin{align*}
    && d(\pi_t) - (b(\pi_t) - c(\pi_t)) &\geq 0 \\ \pause
    \iff && d(\pi) - (b(\pi) - c(\pi))  &\geq 0 \\ \pause
    \iff && d(\pi)                      &\geq b(\pi) - c(\pi).
\end{align*}
\end{proof}

\end{frame}

\begin{frame}

\begin{lemma}[\citeauthor*{Christie1998}]
\label{lem:1}
Let $C$ be the maximum alternating cycle decomposition of $G(\pi)$. \par\pause
Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in $C$ \pause and let $c_{3*}(\pi) = c(\pi) - c_2(\pi)$ be the remaining alternating cycles in $C$. \pause
Then
\begin{align*}
    c_{3*}(\pi) \leq \frac{1}{3}(b(\pi) - 2 c_2(\pi)).
\end{align*}
\end{lemma}

\begin{proof}
\begin{itemize}
    \item $b(\pi)$ is the number of red edges in $G(\pi)$ and $c_2(\pi)$ is the number of red edges in $2$-cycles of $G(\pi)$. \par\pause
    \item $b(\pi) - 2 c_2(\pi)$ is the number of red edges in $G(\pi)$ not occurring in $2$-cycles. \par\pause
    \item $3 c_{3*}(\pi)$ is a lower bound to the number of red edges in $G(\pi)$ not occurring in $2$-cycles.\qedhere
\end{itemize}
\end{proof}

\end{frame}

\begin{frame}

\begin{theorem}[lower bound with $2$-cycles, \citeauthor*{Christie1998}]
\label{thm:3}
Let $\pi \in S_n$. Then
\begin{align*}
    d(\pi) \geq \frac{2}{3} b(\pi) - \frac{1}{3} c_2(\pi).
\end{align*}
\end{theorem}\pause

\begin{proof}
\begin{align*}
    d(\pi) \overset{(\ref{thm:2})}&{\geq} b(\pi) - c(\pi) \\ \pause
           &\geq b(\pi) - c_2(\pi) - c_{3*}(\pi) \\ \pause
           \overset{(\ref{lem:1})}&{\geq} b(\pi) - c_2(\pi) - \frac{1}{3}(b(\pi) - 2 c_2(\pi)) \\ \pause
           &= \frac{2}{3} b(\pi) - \frac{1}{3} c_2(\pi).
\end{align*}
\end{proof}

\end{frame}

\section{3/2-approximation}

\begin{frame}

\begin{block}{$\frac{3}{2}$-approximation}
By theorem \ref{thm:3}, an algorithm that finds a sorting sequence of reversals of at most length $b(\pi) - \frac{1}{2} c_2(\pi)$ achieves an approximation bound of $\frac{3}{2}$. \pause\spadding

We find such an algorithm in two steps:
\begin{enumerate}
    \item given an alternating cycle decomposition $C$ of $G(\pi)$ we find a sorting sequence of reversals for $\pi$\pause; and
    \item we find an alternating cycle decomposition of $G(\pi)$ maximizing the number of $2$-cycles.
\end{enumerate}
\end{block}\pause

Lastly, we prove the approximation bound.

\end{frame}

\subsection{Reversal graph}

\begin{frame}

\begin{definition}[reversal graph of $G(\pi)$, \citeauthor*{Christie1998}]
Given an alternating cycle decomposition $C$ of $G(\pi)$, \par\pause
define $R(C)$ with
\begin{itemize}
    \item an isolated \b{blue} vertex for each adjacency in $\pi$\pause;
    \item $m$ vertices for each $m$-cycle in $C$ each representing the reversal $\rho(u)$ acting on the two red edges connected by a blue edge\pause;
    \begin{itemize}
        \item a vertex is \r{red} if the represented reversal is a $1$- or $2$-reversal
        \item a vertex is \b{blue} otherwise
    \end{itemize}\pause
    \item connect two vertices with an edge if their corresponding blue edges \textit{interleave} (i.e. overlap).
\end{itemize}
\end{definition}\pause

\begin{corollary}[\citeauthor*{Christie1998}]
The only alternating cycle decomposition of $id$ is $C = \emptyset$. \par
$R(C)$ consists of $n+1$ isolated blue vertices.
\end{corollary}

\end{frame}

\begin{frame}

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. Given an alternating cycle decomposition $C$ of $G(\pi)$ construct $R(C)$.\spadding

\begin{tabularx}{\textwidth}{XR}
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture}
with
{\begin{align*}
    C = \{&\{\{0,4\},\{3,4\},\{1,3\},\{0,1\}\} \\
          &\{\{2,4\},\{2,3\},\{3,5\},\{4,5\}\}\}
\end{align*}} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}
\end{exampleblock}\pause

Idea: Each connected component of $R(C)$ can be sorted separately.

\end{frame}

\begin{frame}

Let $u$ be a vertex of $R(C)$.\pause

\begin{definition}
Denote by $C_u$ the alternating cycle decomposition of $G(\pi \circ \rho_u)$ that is obtained from $C$ by removing one 2-cycle or shortening one of its cycles by one red edge.
\end{definition}\pause

\begin{lemma}[\citeauthor*{Christie1998}]
\label{lem:5}
$R(C_u)$ can be derived from $R(C)$ by making the following changes to $R(C)$:\pause
\begin{enumerate}
    \item flip the color of every vertex adjacent to $u$\pause;
    \item flip the adjacency of every pair of vertices adjacent to $u$\pause; and
    \item if $u$ is a red vertex, turn it into an isolated blue vertex.
\end{enumerate}
\end{lemma}

\end{frame}

\begin{frame}

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$ and $u = \{0,1\}$.\padding

\begin{tabularx}{\textwidth}{X|R}
$G(\pi)$ and $R(C)$ & $G(\pi \circ \rho(u))$ and $R(C_u)$ \\ \hline
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {1};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {4};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (4) to (5);
		\draw [style=red edge] (8) to (9);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (5) to (9);
	\end{pgfonlayer}
\end{tikzpicture}\\
\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture} &
\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=blue node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=blue node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=red node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=red node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (2) to (3);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}
\end{exampleblock}

\end{frame}

\begin{frame}

\begin{corollary}[\citeauthor*{Christie1998}]
A reversal $\rho(u)$ affects only vertices that are in the same connected component as $u$.
\end{corollary}\pause

\begin{lemma}[\citeauthor*{Christie1998}]
\label{lem:3}
All vertices arising from the same alternating cycle in $C$ are in the same connected component of $R(C)$.
\end{lemma}

\end{frame}

\begin{frame}

\begin{lemma}[\citeauthor*{Christie1998}]
$R(C)$ contains no isolated blue vertices arising from cycles in $C$.
\end{lemma}\pause

\begin{lemma}[\citeauthor*{Christie1998}]
\label{lem:7}
Vertices arising from an unoriented $2$-cycle of $C$ must be in a connected component of $R(C)$ with vertices arising from at least one more alternating cycle of $C$.
\end{lemma}\pause

\begin{proof}
Let $u$ and $v$ represent the blue edges of an unoriented $2$-cycle of $C$. \par\pause
Then its two constituting blue edges do not interleave. Hence, $R(C)$ does not connect $u$ and $v$ with an edge. \par\pause
If its constituting blue edges do not interleave with blue edges of any other alternating cycle of $C$, then $u$ and $v$ would be isolated blue vertices. \b{$\lightning$}
\end{proof}

\end{frame}

\begin{frame}

\begin{definition}
We call a connected component of $R(C)$ \b{oriented} if it contains a red vertex or if it consists solely of an isolated blue vertex. \pause\spadding

Let $A$ be a connected component of $R(C)$. We denote by $A_u$ the subgraph of $R(C_u)$ that contains all the vertices of $A$.
\end{definition}\pause

\begin{lemma}[\citeauthor*{Christie1998}]
\label{lem:4}
If a connected component $A$ of $R(C)$ is oriented and not an isolated blue vertex, it contains a red vertex $u$ such that $A_u$ is still oriented.
\end{lemma}

\end{frame}

\begin{frame}

% \begin{definition}
% A sequence of reversals that turns the connected component $A$ of $R(C)$ into an isolated blue vertex, is called an \b{elemination sequence} for $A$.
% \end{definition}\pause

\begin{lemma}[\citeauthor*{Christie1998}]
\label{lem:6}
Let $A$ be an oriented connected component of $R(C)$ arising from $k$ different cycles of $G(\pi)$. Then there is an \textit{elimination sequence} for $A$ that contains $k$ $2$-reversals with all the other reversals being $1$-reversals.
\end{lemma}\pause

\begin{proof}
We know that
\begin{itemize}
    \item every vertex in $R(C)$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component (lemma \ref{lem:3})\pause;
    \item  as $A$ is oriented, we can use lemma \ref{lem:4} to repeatedly find a red vertex $u$ in $A$ and apply $\rho(u)$\pause; and
    \item $\rho(u)$ is a $2$-reversal iff $u$ arises from a $2$-cycle (lemma \ref{lem:2}).
\end{itemize}\pause
Every alternating cycle represented by vertices constituting $A$ eventually reduces to a $2$-cycle.
\end{proof}

\end{frame}

\begin{frame}

\begin{lemma}[\citeauthor*{Christie1998}]
Let $A$ be an unoriented connected component of $R(C)$ arising from $k$ different cycles of $G(\pi)$. Then there is an \textit{elimination sequence} for $A$ that contains one $0$-reversal, $k$ $2$-reversals, with all the other reversals being $1$-reversals.
\end{lemma}\pause

\begin{proof}
Let $\rho(u)$ be a $0$-reversal. Then $A_u$ is oriented (lemma \ref{lem:5}). \par\pause
Now apply lemma \ref{lem:6} to obtain an elimination sequence for $A_u$.
\end{proof}\pause

\begin{theorem}[\citeauthor*{Christie1998}]
\label{thm:4}
Let $\pi \in S_n$ and $C$ be a cycle decomposition of $G(\pi)$. Then
\begin{align*}
    d(\pi) \leq b(\pi) - |C| + u(C)
\end{align*}
where $u(C)$ is the number of unoriented components in $R(C)$.
\end{theorem}

\end{frame}

\subsection{Matching graph}

\begin{frame}

Goal: Find a cycle decomposition of $G(\pi)$ that has a large number of $2$-cycles.\pause

\begin{idea}
\begin{enumerate}
    \item Construct a \b{matching graph} $F(\pi)$ where vertices represent red edges in $G(\pi)$ and vertices $u, v$ are adjacent if they form a $2$-cycle in $G(\pi)$. \par\pause
          $\implies$ each edge represents a $2$-cycle.\pause
    \item Find maximum cardinality matching $M$ of $F(\pi)$.\pause
    \item Use a \b{ladder graph} $L(M)$ with vertices representing $2$-cycles in $M$ and form connected components (\textit{ladders}) with $2$-cycles sharing a blue edge in $G(\pi)$.
\end{enumerate}
\end{idea}

\end{frame}

\begin{frame}

\begin{definition}
We call a $2$-cycle \b{selected} if its corresponding edge of $F(\pi)$ is in $M$.\pause\spadding

A selected $2$-cycle is called \b{independent} if it is not part of a ladder. Otherwise it is called a \b{ladder $2$-cycle}.
\end{definition}

\end{frame}

\begin{frame}

Let $z$ be the number of independent $2$-cycles, and $y$ the number of ladder $2$-cycles.\pause

\begin{theorem}[\citeauthor*{Christie1998}]
\label{thm:5}
Given a maximum cardinality matching $M$ of $F(\pi)$ it is possible to find a cycle decomposition $C$ of $G(\pi)$ that contains at least $\left\lceil \frac{y}{2} \right\rceil$ ladder $2$-cycles and $z$ independent $2$-cycles.
\end{theorem}

\begin{proof}
Let $C$ contain all independent $2$-cycles from $L(M)$ and every alternate cycle of each ladder. \par\pause
Obtain the rest of $C$ by adding any alternating cycle decomposition of the remaining unused edges of $G(\pi)$.
\end{proof}

\end{frame}

\subsection{Approximation bound}

\begin{frame}

\begin{theorem}[\citeauthor*{Christie1998}]
\label{thm:6}
Let $\pi \in S_n$. Then
\begin{align*}
    d(\pi) \leq b(\pi) - \frac{1}{2} c_2(\pi).
\end{align*}
\end{theorem}\pause

\begin{proof}\renewcommand{\qedsymbol}{}
Using theorem \ref{thm:5}, first find an alternating cycle decomposition $C$ of $G(\pi)$ with at least $\left\lceil \frac{y}{2} \right\rceil$ $2$-cycles as part of ladders and $z$ independent $2$-cycles.
\end{proof}

\end{frame}

\begin{frame}

\begin{proof}[Proof (cont.)]\renewcommand{\qedsymbol}{}
\begin{itemize}
    \item Let $k$ be the number of $2$-cycles in oriented connected components of $R(C)$.\pause
    \item Let $u$ be the number of unoriented connected components in $R(C)$ that include $l$ selected $2$-cycles and that contain vertices representing remaining unselected $2$-cycles.\pause
    \item Let $v$ be the number of remaining unoriented connected components consisting only of vertices representing $m$ independent selected $2$-cycles.
\end{itemize}\pause
By theorem \ref{thm:4}, we can sort $\pi$ using at least $k+l+u+m$ $2$-reversals and only $u+v$ $0$-reversals. \pause Therefore
\begin{align*}
    d(\pi) &\leq b(\pi) - k - l - u - m + u + v \\
           &= b(\pi) - k - l - m + v
\end{align*}
\end{proof}

\end{frame}

\begin{frame}

\begin{proof}[Proof (cont.)]
We know that
\begin{enumerate}
    \item $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ as $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles by $M$\pause;
    \item $v \leq \left\lceil \frac{z}{2} \right\rceil$ as every unoriented component representing a $2$-cycle represents at least one more alternating cycle (lemma \ref{lem:7})\pause; and
    \item $|M| = y+z \geq c_2(\pi)$.
\end{enumerate}\pause
\begin{align*}
    k+l+m-v &\geq \left\lceil \frac{y}{2} \right\rceil + z - v                                          & (1) \\ \pause
            &\geq \left\lceil \frac{y}{2} \right\rceil + z - \left\lceil \frac{z}{2} \right\rceil       & (2) \\ \pause
            &= \left\lceil \frac{y}{2} \right\rceil + \left\lceil \frac{z}{2} \right\rceil \\ \pause
            &\geq \frac{c_2(\pi)}{2}                                                                    & (3)
\end{align*}
\end{proof}

\end{frame}

\begin{frame}[allowframebreaks]{References}

\printbibliography

\end{frame}

\end{document}

### 2020-06-29 slides.tex
improvements

@@
 
-\usepackage{biblatex}
+\usepackage[sorting=ynt]{biblatex}
 \addbibresource{sources.bib}
@@
 
-\begin{algorithm}[Watterson]
-For each element at position $i$, perform $\rho(i, \pi_i^{-1})$ if $i \neq \pi_i$. \cite{Kececioglu1995}
-\end{algorithm}\pause\padding
+\begin{algorithm}[ratchet algorithm, \citeauthor*{Watterson19821}]
+For each element at position $i$, perform $\rho(i, \pi_i^{-1})$ if $i \neq \pi_i$.
+\end{algorithm}\pause
 
@@
 
-An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges.
+An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges. \pause\spadding
+
+We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$.
 \end{definition}
-
-\end{frame}
-
-\begin{frame}
-
-\begin{definition}
-We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$. \pause\spadding
-
-An alternating cycle in $G(\pi)$ is \b{directed} with respect to $\pi$ if it is possible to walk along the whole cycle traversing each red edge in the direction of its orientation.
-\end{definition}
-
-\end{frame}
-
-\begin{frame}
-
-\begin{proposition}[\citeauthor*{Caprara1997}]
-An alternating cycle in $G(\pi)$ is oriented iff it is undirected.
-\end{proposition}\pause
-
-\begin{exampleblock}{Example}
-\begin{center}
-\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=node] (6) at (-3.75, 0) {0};
-		\node [style=node] (7) at (-2.25, 0) {3};
-		\node [style=node] (8) at (-0.75, 0) {4};
-		\node [style=node] (9) at (0.75, 0) {1};
-		\node [style=node] (10) at (2.25, 0) {2};
-		\node [style=node] (11) at (3.75, 0) {5};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw [style=red edge] (6) to (7);
-		\draw [style=red edge] (8) to (9);
-		\draw [style=red edge] (10) to (11);
-		\draw [style=blue edge, bend right] (6) to (9);
-		\draw [style=blue edge, bend left] (7) to (10);
-		\draw [style=blue edge, bend right] (8) to (11);
-	\end{pgfonlayer}
-\end{tikzpicture}
-\end{center}\pause
-$\implies$ directed and unoriented.
-\end{exampleblock}
 
@@
 \begin{exampleblock}{Example}
-Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. Given an alternating cycle decomposition $C$ of $G(\pi)$ construct $R(C)$.\spadding
+Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. \par
+Given the alternating cycle decomposition $C$ of $G(\pi)$
 
 \begin{tabularx}{\textwidth}{XR}
+{\footnotesize\begin{align*}
+    C = \{&\{\{0,4\},\{3,4\},\{1,3\},\{0,1\}\} \\
+          &\{\{2,4\},\{2,3\},\{3,5\},\{4,5\}\}\}
+\end{align*}} &
 \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
@@
 \end{tikzpicture}
-with
-{\begin{align*}
-    C = \{&\{\{0,4\},\{3,4\},\{1,3\},\{0,1\}\} \\
-          &\{\{2,4\},\{2,3\},\{3,5\},\{4,5\}\}\}
-\end{align*}} &
-\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+\end{tabularx}\pause
+
+construct $R(C)$.
+
+\begin{center}
+\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
 	\begin{pgfonlayer}{nodelayer}
@@
 \end{tikzpicture}
-\end{tabularx}
+\end{center}
 \end{exampleblock}\pause
@@
 \end{tikzpicture}\\
-\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
 	\begin{pgfonlayer}{nodelayer}
@@
 \end{tikzpicture} &
-\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
 	\begin{pgfonlayer}{nodelayer}

### 2020-06-30 slides.tex
improvements

@@
 and $d$ is minimal. \pause\par\padding
-$d$ is called \b{reversal distance} between $\sigma$ and $\tau$. \cite{Kececioglu1995}
+$d$ is called \b{reversal distance} between $\sigma$ and $\tau$.
 \end{definition}
@@
 \end{align*}
-and $d$ is minimal. \cite{Kececioglu1995}
+and $d$ is minimal.
 \end{definition}
@@
 $\implies d(\pi) \leq 2$.
-\end{exampleblock}\pause\padding
+\end{exampleblock}
 
-\begin{algorithm}[ratchet algorithm, \citeauthor*{Watterson19821}]
-For each element at position $i$, perform $\rho(i, \pi_i^{-1})$ if $i \neq \pi_i$.
-\end{algorithm}\pause
+% \begin{algorithm}[ratchet algorithm, \citeauthor*{Watterson19821}]
+% For each element at position $i$, perform $\rho(i, \pi_i^{-1})$ if $i \neq \pi_i$.
+% \end{algorithm}\pause
 
-\begin{corollary}[tight upper bound, \citeauthor*{Kececioglu1995}]
-$d(\pi) \leq n - 1$ for all $\pi \in S_n$.
-\end{corollary}
+% \begin{corollary}[tight upper bound, \citeauthor*{Kececioglu1995}]
+% $d(\pi) \leq n - 1$ for all $\pi \in S_n$.
+% \end{corollary}
 
@@
 
-Observation: $b(\pi) = 0$ iff $\pi = id$ and any reversal can at most eliminate two breakpoints. \pause\spadding
+Observation: $b(\pi) = 0$ iff $\pi = id$ \pause and any reversal can at most eliminate two breakpoints. \pause\spadding
 
@@
         \item a \r{red} edge for every breakpoint in $\pi$; and
-        \item a \b{blue} edge for every adjacency in $\pi$.
+        \item a \b{blue} edge for every missing adjacency in $\pi$.
     \end{itemize}
@@
 \begin{exampleblock}{Example}
-Let $\pi = (0\ \r\mid\ 4\ \r\mid\ 2\ 1\ \r\mid\ 3\ \r\mid\ 5) \in S_4$.
+Let $\pi = (0\ \r\mid\ 4\ \r\mid\ 2\ 1\ \r\mid\ 3\ \r\mid\ 5) \in S_4$. \pause Then $G(\pi)$ is
 
@@
 
+\begin{center}
+\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node] (0) at (-3, 0) {0};
+		\node [style=node] (3) at (-1.5, 0) {4};
+		\node [style=node] (4) at (0, 0) {2};
+		\node [style=node] (5) at (1.5, 0) {1};
+		\node [style=node] (8) at (3, 0) {3};
+		\node [style=node] (9) at (4.5, 0) {5};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=red edge] (3) to (4);
+		\draw [style=red edge] (8) to (9);
+		\draw [style=red edge] (0) to (3);
+		\draw [style=red edge] (5) to (8);
+		\draw [style=blue edge, bend left] (4) to (8);
+		\draw [style=blue edge, bend left] (3) to (9);
+		\draw [style=blue edge, bend right] (0) to (5);
+		\draw [style=blue edge, bend right] (3) to (8);
+	\end{pgfonlayer}
+\end{tikzpicture}
+\end{center}
+
 \begin{definition}
@@
 
-A reversal \b{acts on} two red edges of $G\pi)$ if those two edges represent the breakpoints that are split apart by the reversal. \pause\spadding
+A reversal \b{acts on} two red edges of $G(\pi)$ if those two edges represent the breakpoints that are split apart by the reversal. \pause\spadding
 
@@
 
-\begin{definition}[reversal graph of $G(\pi)$, \citeauthor*{Christie1998}]
+\begin{definition}[reversal graph, \citeauthor*{Christie1998}]
 Given an alternating cycle decomposition $C$ of $G(\pi)$, \par\pause
@@
 \begin{definition}
-Denote by $C_u$ the alternating cycle decomposition of $G(\pi \circ \rho_u)$ that is obtained from $C$ by removing one 2-cycle or shortening one of its cycles by one red edge.
+Denote by $C_u$ the alternating cycle decomposition of $G(\pi \circ \rho(u))$ that is obtained from $C$ by removing one 2-cycle or shortening one of its cycles by one red edge.
 \end{definition}\pause

### 2020-07-03 slides.tex
update examples

@@
 
-\usepackage[sorting=ynt]{biblatex}
+\usepackage[sorting=ynt,style=alphabetic]{biblatex}
 \addbibresource{sources.bib}
@@
 
-\section{Introduction}
+\section{Motivation}
 
@@
 \begin{exampleblock}{Example}
-Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. \par
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$. \par
 Then
 \begin{align*}
-    \pi \circ \rho(1,4) = (0\ \r{3\ 1\ 2\ 4}\ 5).
+    \pi \circ \rho(2,4) = (0\ 1\ \r{2\ 4\ 3}\ 5).
 \end{align*}
@@
 \begin{exampleblock}{Example}
-Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$.\pause
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$.\pause
 \begin{align*}
-    \pi \circ \rho(1,3) &= (0\ \r{1\ 2\ 4}\ 3\ 5) \\ \pause
-    \pi \circ \rho(1,3) \circ \rho(3,4) &= (0\ 1\ 2\ \r{3\ 4}\ 5) = id
+    \pi \circ \rho(2,4) &= (0\ 1\ \r{2\ 4\ 3}\ 5) \\ \pause
+    \pi \circ \rho(2,4) \circ \rho(3,4) &= (0\ 1\ 2\ \r{3\ 4}\ 5) = id
 \end{align*}\pause
@@
 \begin{exampleblock}{Example}
-Let $\pi = (0\ \r\mid\ 4\ \r\mid\ 2\ 1\ \r\mid\ 3\ \r\mid\ 5) \in S_4$. \pause Then $G(\pi)$ is
+Let $\pi = (0\ 1\ \r\mid\ 3\ 4\ \r\mid\ 2\ \r\mid\ 5) \in S_4$. \pause Then $G(\pi)$ is
 
-\begin{center}
+\spadding\begin{center}
 \begin{tikzpicture}
@@
 		\node [style=node] (0) at (-3, 0) {0};
-		\node [style=node] (3) at (-1.5, 0) {4};
-		\node [style=node] (4) at (0, 0) {2};
-		\node [style=node] (5) at (1.5, 0) {1};
-		\node [style=node] (8) at (3, 0) {3};
-		\node [style=node] (9) at (4.5, 0) {5};
+		\node [style=node] (1) at (-1.5, 0) {1};
+		\node [style=node] (2) at (0, 0) {3};
+		\node [style=node] (3) at (1.5, 0) {4};
+		\node [style=node] (4) at (3, 0) {2};
+		\node [style=node] (5) at (4.5, 0) {5};
 	\end{pgfonlayer}
 	\begin{pgfonlayer}{edgelayer}
+		\draw [style=red edge] (1) to (2);
 		\draw [style=red edge] (3) to (4);
-		\draw [style=red edge] (8) to (9);
-		\draw [style=red edge] (0) to (3);
-		\draw [style=red edge] (5) to (8);
-		\draw [style=blue edge, bend left] (4) to (8);
-		\draw [style=blue edge, bend left] (3) to (9);
-		\draw [style=blue edge, bend right] (0) to (5);
-		\draw [style=blue edge, bend right] (3) to (8);
+		\draw [style=red edge] (4) to (5);
+		\draw [style=blue edge, bend right] (1) to (4);
+		\draw [style=blue edge, bend left] (2) to (4);
+		\draw [style=blue edge, bend left] (3) to (5);
 	\end{pgfonlayer}
@@
     \end{itemize}\pause
-    \item connect two vertices with an edge if their corresponding blue edges \textit{interleave} (i.e. overlap).
+    \item connect two vertices with an edge if their corresponding blue edges \textit{interleave}.
 \end{itemize}
@@
 \begin{corollary}[\citeauthor*{Christie1998}]
-The only alternating cycle decomposition of $id$ is $C = \emptyset$. \par
+The only alternating cycle decomposition of $G(id)$ is $C = \emptyset$. \par
 $R(C)$ consists of $n+1$ isolated blue vertices.
@@
 {\footnotesize\begin{align*}
-    C = \{&\{\{0,4\},\{3,4\},\{1,3\},\{0,1\}\} \\
-          &\{\{2,4\},\{2,3\},\{3,5\},\{4,5\}\}\}
+    C = \{(&\{1,3\},\{2,3\},\{2,4\}, \\
+           &\{4,5\},\{2,5\},\{1,2\})\}
 \end{align*}} &
@@
 	\begin{pgfonlayer}{nodelayer}
-		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
-		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
-		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
-		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
-		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
+		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
+		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
+		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
+		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
 	\end{pgfonlayer}
 	\begin{pgfonlayer}{edgelayer}
-		\draw (1) to (0);
-		\draw (2) to (1);
-		\draw (0) to (3);
-		\draw (1) to (3);
-		\draw (2) to (0);
+		\draw (0) to (2);
+		\draw (1) to (2);
 	\end{pgfonlayer}
@@
 \begin{exampleblock}{Example}
-Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$ and $u = \{0,1\}$.\padding
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$ and $u = \{1,2\}$.\padding
 
@@
 $G(\pi)$ and $R(C)$ & $G(\pi \circ \rho(u))$ and $R(C_u)$ \\ \hline
-\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
 	\begin{pgfonlayer}{nodelayer}
 		\node [style=node] (0) at (-3, 0) {0};
-		\node [style=node] (3) at (-1.5, 0) {4};
-		\node [style=node] (4) at (0, 0) {2};
-		\node [style=node] (5) at (1.5, 0) {1};
-		\node [style=node] (8) at (3, 0) {3};
-		\node [style=node] (9) at (4.5, 0) {5};
+		\node [style=node] (1) at (-1.5, 0) {1};
+		\node [style=node] (2) at (0, 0) {3};
+		\node [style=node] (3) at (1.5, 0) {4};
+		\node [style=node] (4) at (3, 0) {2};
+		\node [style=node] (5) at (4.5, 0) {5};
 	\end{pgfonlayer}
 	\begin{pgfonlayer}{edgelayer}
+		\draw [style=red edge] (1) to (2);
 		\draw [style=red edge] (3) to (4);
-		\draw [style=red edge] (8) to (9);
-		\draw [style=red edge] (0) to (3);
-		\draw [style=red edge] (5) to (8);
-		\draw [style=blue edge, bend left] (4) to (8);
-		\draw [style=blue edge, bend left] (3) to (9);
-		\draw [style=blue edge, bend right] (0) to (5);
-		\draw [style=blue edge, bend right] (3) to (8);
+		\draw [style=red edge] (4) to (5);
+		\draw [style=blue edge, bend right] (1) to (4);
+		\draw [style=blue edge, bend left] (2) to (4);
+		\draw [style=blue edge, bend left] (3) to (5);
 	\end{pgfonlayer}
@@
 		\node [style=node] (0) at (-3, 0) {0};
-		\node [style=node] (3) at (-1.5, 0) {1};
-		\node [style=node] (4) at (0, 0) {2};
-		\node [style=node] (5) at (1.5, 0) {4};
-		\node [style=node] (8) at (3, 0) {3};
-		\node [style=node] (9) at (4.5, 0) {5};
+		\node [style=node] (1) at (-1.5, 0) {1};
+		\node [style=node] (2) at (0, 0) {2};
+		\node [style=node] (3) at (1.5, 0) {4};
+		\node [style=node] (4) at (3, 0) {3};
+		\node [style=node] (5) at (4.5, 0) {5};
 	\end{pgfonlayer}
 	\begin{pgfonlayer}{edgelayer}
+		\draw [style=red edge] (2) to (3);
 		\draw [style=red edge] (4) to (5);
-		\draw [style=red edge] (8) to (9);
-		\draw [style=blue edge, bend left] (4) to (8);
-		\draw [style=blue edge, bend left] (5) to (9);
+		\draw [style=blue edge, bend left] (2) to (4);
+		\draw [style=blue edge, bend left] (3) to (5);
 	\end{pgfonlayer}
@@
 	\begin{pgfonlayer}{nodelayer}
-		\node [style=blue node] (0) at (0, 1) {$\{3,4\}$};
-		\node [style=blue node] (1) at (-2, 1) {$\{0,1\}$};
-		\node [style=red node] (2) at (-2, -1) {$\{2,3\}$};
-		\node [style=red node] (3) at (0, -1) {$\{4,5\}$};
-		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
+		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
+		\node [style=red node] (2) at (-1, -1) {$\{4,5\}$};
+		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
+		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
 	\end{pgfonlayer}
 	\begin{pgfonlayer}{edgelayer}
-		\draw (2) to (3);
+		\draw (0) to (2);
 	\end{pgfonlayer}

### 2020-07-04 slides.tex
update slides

@@
 \begin{proof}
+To show: $b(\pi) - b(\pi \circ \rho) + c(\pi \circ \rho) - c(\pi) \leq 1$. \par
 We consider each case $b(\pi) - b(\pi \circ \rho) \in [-2, 2]$ separately.\pause
@@
 
+Observation: The only alternating cycle decomposition of $G(id)$ is $C = \emptyset$.\pause
+
 \begin{corollary}[\citeauthor*{Christie1998}]
-The only alternating cycle decomposition of $G(id)$ is $C = \emptyset$. \par
 $R(C)$ consists of $n+1$ isolated blue vertices.
@@
 \end{tikzpicture} &
-\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+\onslide<2->{\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}\\
+\end{tikzpicture}}\\
 \spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
@@
 \end{tikzpicture} &
-\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
+\onslide<3->{\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}
+\end{tikzpicture}}
 \end{tabularx}
@@
 
+\begin{corollary}[\citeauthor*{Christie1998}]
+$R(C)$ contains no isolated blue vertices arising from cycles in $C$.
+\end{corollary}\pause
+
 \end{frame}
@@
 \begin{lemma}[\citeauthor*{Christie1998}]
-$R(C)$ contains no isolated blue vertices arising from cycles in $C$.
-\end{lemma}\pause
-
-\begin{lemma}[\citeauthor*{Christie1998}]
 \label{lem:7}
 Vertices arising from an unoriented $2$-cycle of $C$ must be in a connected component of $R(C)$ with vertices arising from at least one more alternating cycle of $C$.
-\end{lemma}\pause
+\end{lemma}
+
+\begin{center}
+\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node] (0) at (-3, 0) {};
+		\node [style=node] (1) at (-1.25, 0) {};
+		\node [style=node] (2) at (1.25, 0) {};
+		\node [style=node] (3) at (3, 0) {};
+		\node (4) at (0, 0) {$\dots$};
+		\node (5) at (-4.25, 0) {$\dots$};
+		\node (6) at (4.25, 0) {$\dots$};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=blue edge, bend left] (0) to node[sloped,midway,above] {u} (3);
+		\draw  [style=blue edge, bend right](1) to node[sloped,midway,below] {v} (2);
+		\draw [style=red edge] (0) to (1);
+		\draw [style=red edge] (2) to (3);
+	\end{pgfonlayer}
+\end{tikzpicture}
+\end{center}\pause
 

### 2020-07-05 slides.tex
update slides

@@
 \begin{proof}\renewcommand{\qedsymbol}{}
-Let $\pi_t = \pi, \pi_0 = id$ and $\rho_1, \dots, \rho_t$ a shortest sequence of reversals from $\pi_t$ to $\pi_0$. \par\pause
-Then
+Let $\pi_t = \pi, \pi_0 = id$ and $\rho_1, \dots, \rho_t$ a shortest sequence of reversals from $\pi_t$ to $\pi_0$. \pause Then
 \begin{align*}
     d(\pi_i) &= d(\pi_{i-1}) + 1 \\ \pause
-             \overset{(\ref{thm:1})}&{\geq} d(\pi_{i-1}) + b(\pi_i) - b(\pi_{i-1}) + c(\pi_{i-1}) - c(\pi_i).
+             \overset{(\ref{thm:1})}&{\geq} d(\pi_{i-1}) + b(\pi_i) - b(\pi_{i-1}) + c(\pi_{i-1}) - c(\pi_i) \\[0.25cm] \pause
+    \iff & \begin{aligned}[t]
+    d(\pi_i) - (b(\pi_i) - c(\pi_i)) &\geq d(\pi_{i-1}) - (b(\pi_{i-1}) - c(\pi_{i-1})) \\ \pause
+                                     &\geq d(\pi_0) - (b(\pi_0) - c(\pi_0))\pause = 0
+    \end{aligned}
 \end{align*}
@@
 \begin{proof}[Proof (cont.)]
-Therefore
-\begin{align*}
-    d(\pi_i) - (b(\pi_i) - c(\pi_i)) &\geq d(\pi_{i-1}) - (b(\pi_{i-1}) - c(\pi_{i-1})) \\ \pause
-                                     &\geq d(\pi_0) - (b(\pi_0) - c(\pi_0))\pause = 0
-\end{align*}\pause
 Let $i = t$. Then
 \begin{align*}
-    && d(\pi_t) - (b(\pi_t) - c(\pi_t)) &\geq 0 \\ \pause
-    \iff && d(\pi) - (b(\pi) - c(\pi))  &\geq 0 \\ \pause
+         && d(\pi) - (b(\pi) - c(\pi)) &\geq 0 \\ \pause
     \iff && d(\pi)                     &\geq b(\pi) - c(\pi).
 \end{align*}
-\end{proof}
-
-\end{frame}
-
-\begin{frame}
-
-\begin{lemma}[\citeauthor*{Christie1998}]
-\label{lem:1}
-Let $C$ be the maximum alternating cycle decomposition of $G(\pi)$. \par\pause
-Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in $C$ \pause and let $c_{3*}(\pi) = c(\pi) - c_2(\pi)$ be the remaining alternating cycles in $C$. \pause
-Then
-\begin{align*}
-    c_{3*}(\pi) \leq \frac{1}{3}(b(\pi) - 2 c_2(\pi)).
-\end{align*}
-\end{lemma}
-
-\begin{proof}
-\begin{itemize}
-    \item $b(\pi)$ is the number of red edges in $G(\pi)$ and $c_2(\pi)$ is the number of red edges in $2$-cycles of $G(\pi)$. \par\pause
-    \item $b(\pi) - 2 c_2(\pi)$ is the number of red edges in $G(\pi)$ not occurring in $2$-cycles. \par\pause
-    \item $3 c_{3*}(\pi)$ is a lower bound to the number of red edges in $G(\pi)$ not occurring in $2$-cycles.\qedhere
-\end{itemize}
-\end{proof}
-
-\end{frame}
-
-\begin{frame}
+\end{proof}\pause
 
@@
 \label{thm:3}
-Let $\pi \in S_n$. Then
+Let $\pi \in S_n$ and $C$ be the maximum alternating cycle decomposition of $G(\pi)$. Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in $C$. Then
 \begin{align*}
@@
 \end{align*}
-\end{theorem}\pause
-
-\begin{proof}
-\begin{align*}
-    d(\pi) \overset{(\ref{thm:2})}&{\geq} b(\pi) - c(\pi) \\ \pause
-           &\geq b(\pi) - c_2(\pi) - c_{3*}(\pi) \\ \pause
-           \overset{(\ref{lem:1})}&{\geq} b(\pi) - c_2(\pi) - \frac{1}{3}(b(\pi) - 2 c_2(\pi)) \\ \pause
-           &= \frac{2}{3} b(\pi) - \frac{1}{3} c_2(\pi).
-\end{align*}
-\end{proof}
+\end{theorem}
 
@@
 \begin{exampleblock}{Example}
-Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. \par
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$. \par
 Given the alternating cycle decomposition $C$ of $G(\pi)$
@@
 All vertices arising from the same alternating cycle in $C$ are in the same connected component of $R(C)$.
-\end{lemma}
+\end{lemma}\pause
 
@@
 $R(C)$ contains no isolated blue vertices arising from cycles in $C$.
-\end{corollary}\pause
+\end{corollary}
 
@@
     \item Construct a \b{matching graph} $F(\pi)$ where vertices represent red edges in $G(\pi)$ and vertices $u, v$ are adjacent if they form a $2$-cycle in $G(\pi)$. \par\pause
-          $\implies$ each edge represents a $2$-cycle.\pause
     \item Find maximum cardinality matching $M$ of $F(\pi)$.\pause

### 2020-07-07 slides.tex
update slides

@@
 \]
-for some $j \geq i$.
+for some $i, j \in [1, n]$ with $j \geq i$.
 \end{definition}\pause
@@
 An oriented alternating cycle is a $2$-cycle iff the corresponding reversal acting on both red edges is a $2$-reversal.
-\end{lemma}\pause
+\end{lemma}\spadding
 
-\begin{proof}
 Let $i \sim i'$ and $j \sim j'$.
 
+\begin{figure}
 \begin{center}
@@
 \end{center}
-\end{proof}
+\caption{Oriented $2$-cycle}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node] (0) at (-1.25, 0) {i};
+		\node [style=node] (1) at (1.25, 0) {i'};
+		\node [style=node] (2) at (3, 0) {j'};
+		\node [style=node] (3) at (-3, 0) {j};
+		\node (4) at (0, 0) {$\dots$};
+		\node (5) at (-4.25, 0) {$\dots$};
+		\node (6) at (4.25, 0) {$\dots$};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=blue edge, bend right] (2) to (3);
+		\draw [style=blue edge, bend left] (0) to (1);
+		\draw [style=red edge] (3) to (0);
+		\draw [style=red edge] (1) to (2);
+	\end{pgfonlayer}
+\end{tikzpicture}
+\end{center}
+\caption{Unoriented $2$-cycle}
+\end{figure}
 
@@
 \label{thm:3}
-Let $\pi \in S_n$ and $C$ be the maximum alternating cycle decomposition of $G(\pi)$. Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in $C$. Then
+Let $\pi \in S_n$ and $C$ be a maximum alternating cycle decomposition of $G(\pi)$. Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in any such $C$. Then
 \begin{align*}
@@
 
-Let $u$ be a vertex of $R(C)$.\pause
+Let $u$ be a vertex of $R(C)$ representing reversal $\rho(u)$.\pause
 
 \begin{definition}
-Denote by $C_u$ the alternating cycle decomposition of $G(\pi \circ \rho(u))$ that is obtained from $C$ by removing one 2-cycle or shortening one of its cycles by one red edge.
+Denote by $C_u$ the alternating cycle decomposition of $G(\pi \circ \rho(u))$ that is obtained from $C$.% by removing one 2-cycle or shortening one of its cycles by one red edge.
 \end{definition}\pause
@@
 
-\begin{corollary}[\citeauthor*{Christie1998}]
-A reversal $\rho(u)$ affects only vertices that are in the same connected component as $u$.
-\end{corollary}\pause
+% \begin{corollary}[\citeauthor*{Christie1998}]
+% A reversal $\rho(u)$ affects only vertices that are in the same connected component as $u$.
+% \end{corollary}\pause
 
-\begin{lemma}[\citeauthor*{Christie1998}]
-\label{lem:3}
-All vertices arising from the same alternating cycle in $C$ are in the same connected component of $R(C)$.
-\end{lemma}\pause
-
-\begin{corollary}[\citeauthor*{Christie1998}]
-$R(C)$ contains no isolated blue vertices arising from cycles in $C$.
-\end{corollary}
+% \begin{corollary}[\citeauthor*{Christie1998}]
+% $R(C)$ contains no isolated blue vertices arising from cycles in $C$.
+% \end{corollary}
 
@@
 
+\begin{lemma}[\citeauthor*{Christie1998}]
+\label{lem:3}
+All vertices arising from the same alternating cycle in $C$ are in the same connected component of $R(C)$.
+\end{lemma}\pause
+
 \begin{definition}
@@
 
-\end{frame}
-
-\begin{frame}
-
 % \begin{definition}
@@
 % \end{definition}\pause
-
-\begin{lemma}[\citeauthor*{Christie1998}]
-\label{lem:6}
-Let $A$ be an oriented connected component of $R(C)$ arising from $k$ different cycles of $G(\pi)$.
-Then there is an \textit{elimination sequence} for $A$ that contains $k$ $2$-reversals with all the other reversals being $1$-reversals.
-\end{lemma}\pause
-
-\begin{proof}
-We know that
-\begin{itemize}
-    \item every vertex in $R(C)$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component (lemma \ref{lem:3})\pause;
-    \item  as $A$ is oriented, we can use lemma \ref{lem:4} to repeatedly find a red vertex $u$ in $A$ and apply $\rho(u)$\pause; and
-    \item $\rho(u)$ is a $2$-reversal iff $u$ arises from a $2$-cycle (lemma \ref{lem:2}).
-\end{itemize}\pause
-Every alternating cycle represented by vertices constituting $A$ eventually reduces to a $2$-cycle.
-\end{proof}
 
@@
 
-\begin{lemma}[\citeauthor*{Christie1998}]
-Let $A$ be an unoriented connected component of $R(C)$ arising from $k$ different cycles of $G(\pi)$.
+\begin{exampleblock}{Example (elimination sequence)}
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$\onslide<2->{ and $u_1 = \{1,2\}$}\onslide<3->{, $u_2 = \{2,3\}$.}\spadding
+
+\only<1>{\begin{center}\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node] (0) at (-3, 0) {0};
+		\node [style=node] (1) at (-1.5, 0) {1};
+		\node [style=node] (2) at (0, 0) {3};
+		\node [style=node] (3) at (1.5, 0) {4};
+		\node [style=node] (4) at (3, 0) {2};
+		\node [style=node] (5) at (4.5, 0) {5};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=red edge] (1) to (2);
+		\draw [style=red edge] (3) to (4);
+		\draw [style=red edge] (4) to (5);
+		\draw [style=blue edge, bend right] (1) to (4);
+		\draw [style=blue edge, bend left] (2) to (4);
+		\draw [style=blue edge, bend left] (3) to (5);
+	\end{pgfonlayer}
+\end{tikzpicture}\end{center}}\spadding
+\only<1>{\begin{center}\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
+		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
+		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
+		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw (0) to (2);
+		\draw (1) to (2);
+	\end{pgfonlayer}
+\end{tikzpicture}\end{center}}
+\only<2>{\begin{center}\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node] (0) at (-3, 0) {0};
+		\node [style=node] (1) at (-1.5, 0) {1};
+		\node [style=node] (2) at (0, 0) {2};
+		\node [style=node] (3) at (1.5, 0) {4};
+		\node [style=node] (4) at (3, 0) {3};
+		\node [style=node] (5) at (4.5, 0) {5};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=red edge] (2) to (3);
+		\draw [style=red edge] (4) to (5);
+		\draw [style=blue edge, bend left] (2) to (4);
+		\draw [style=blue edge, bend left] (3) to (5);
+	\end{pgfonlayer}
+\end{tikzpicture}\end{center}}\spadding
+\only<2>{\begin{center}\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
+		\node [style=red node] (2) at (-1, -1) {$\{4,5\}$};
+		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
+		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw (0) to (2);
+	\end{pgfonlayer}
+\end{tikzpicture}\end{center}}
+\only<3>{\begin{center}\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node] (0) at (-3, 0) {0};
+		\node [style=node] (1) at (-1.5, 0) {1};
+		\node [style=node] (2) at (0, 0) {2};
+		\node [style=node] (3) at (1.5, 0) {3};
+		\node [style=node] (4) at (3, 0) {4};
+		\node [style=node] (5) at (4.5, 0) {5};
+	\end{pgfonlayer}
+\end{tikzpicture}\end{center}}\spadding
+\only<3>{\begin{center}\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=blue node] (0) at (0, 1) {$\{2,3\}$};
+		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
+		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
+		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
+		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
+	\end{pgfonlayer}
+\end{tikzpicture}\end{center}}
+\end{exampleblock}
+
+\end{frame}
+
+% \begin{frame}
+
+% \begin{lemma}[\citeauthor*{Christie1998}]
+% \label{lem:6}
+% Let $A$ be an oriented connected component of $R(C)$ arising from $k$ different alternating cycles of $G(\pi)$.
+Then there is an \textit{elimination sequence} for $A$ that contains $k$ $2$-reversals with all the other reversals being $1$-reversals.
+% \end{lemma}\pause
+
+% \begin{proof}
+% We know that
+% \begin{itemize}
+%     \item every vertex in $R(C)$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component (lemma \ref{lem:3})\pause;
+%     \item  as $A$ is oriented, we can use lemma \ref{lem:4} to repeatedly find a red vertex $u$ in $A$ and apply $\rho(u)$\pause; and
+%     \item $\rho(u)$ is a $2$-reversal iff $u$ arises from a $2$-cycle (lemma \ref{lem:2}).
+% \end{itemize}\pause
+% Every alternating cycle represented by vertices constituting $A$ eventually reduces to a $2$-cycle.
+% \end{proof}
+
+% \end{frame}
+
+\begin{frame}
+
+% \begin{lemma}[\citeauthor*{Christie1998}]
+% Let $A$ be an unoriented connected component of $R(C)$ arising from $k$ different alternating cycles of $G(\pi)$.
 Then there is an \textit{elimination sequence} for $A$ that contains one $0$-reversal, $k$ $2$-reversals, with all the other reversals being $1$-reversals.
-\end{lemma}\pause
+% \end{lemma}\pause
 
-\begin{proof}
-Let $\rho(u)$ be a $0$-reversal. Then $A_u$ is oriented (lemma \ref{lem:5}). \par\pause
-Now apply lemma \ref{lem:6} to obtain an elimination sequence for $A_u$.
-\end{proof}\pause
+% \begin{proof}
+% Let $\rho(u)$ be a $0$-reversal. Then $A_u$ is oriented (lemma \ref{lem:5}). \par\pause
+% Now apply lemma \ref{lem:6} to obtain an elimination sequence for $A_u$.
+% \end{proof}\pause
+
+Observations:
+\begin{itemize}
+    \item Every connected component $A$ arising from k different alternating cycles of $G(\pi)$, eventually reduces to a two cycle.\pause
+    \item There exists an elimination sequence of $A$ with $k$ $2$-reversals and remaining $1$-reversals.\pause
+    \item An unoriented connected component requires one initial $0$-reversal.
+\end{itemize}\spadding\pause
 
@@
 A selected $2$-cycle is called \b{independent} if it is not part of a ladder. Otherwise it is called a \b{ladder $2$-cycle}.
-\end{definition}
-
-\end{frame}
-
-\begin{frame}
+\end{definition}\pause\padding
 
@@
 
-\begin{proof}
-Let $C$ contain all independent $2$-cycles from $L(M)$ and every alternate cycle of each ladder. \par\pause
-Obtain the rest of $C$ by adding any alternating cycle decomposition of the remaining unused edges of $G(\pi)$.
-\end{proof}
+% \begin{proof}
+% Let $C$ contain all independent $2$-cycles from $L(M)$ and every alternate cycle of each ladder. \par\pause
+% Obtain the rest of $C$ by adding any alternating cycle decomposition of the remaining unused edges of $G(\pi)$.
+% \end{proof}
 
@@
 \begin{proof}[Proof (cont.)]
-We know that
+To show: $k+l+m-v \geq \frac{c_2(\pi)}{2}$. \pause We know that
 \begin{enumerate}
     \item $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ as $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles by $M$\pause;
-    \item $v \leq \left\lceil \frac{z}{2} \right\rceil$ as every unoriented component representing a $2$-cycle represents at least one more alternating cycle (lemma \ref{lem:7})\pause; and
+    \item $v \leq \left\lfloor \frac{z}{2} \right\rfloor$ as every unoriented component representing a $2$-cycle represents at least one more alternating cycle (lemma \ref{lem:7})\pause; and
     \item $|M| = y+z \geq c_2(\pi)$.
@@
     k+l+m-v &\geq \left\lceil \frac{y}{2} \right\rceil + z - v                                          & (1) \\ \pause
-            &\geq \left\lceil \frac{y}{2} \right\rceil + z - \left\lceil \frac{z}{2} \right\rceil       & (2) \\ \pause
+            &\geq \left\lceil \frac{y}{2} \right\rceil + z - \left\lfloor \frac{z}{2} \right\rfloor     & (2) \\ \pause
             &= \left\lceil \frac{y}{2} \right\rceil + \left\lceil \frac{z}{2} \right\rceil \\ \pause
@@
 
+\begin{frame}
+
+Run time: $O(n^4)$, can be improved to $O(n^2)$.\spadding
+
+\begin{block}{Summary}
+\begin{itemize}
+    \item the number of alternating cycles in a breakpoint graph $G(\pi)$ are related to $d(\pi)$
+    \item a sorting sequence of reversals can be constructed from an alternating cycle decomposition of $G(\pi)$
+\end{itemize}
+\end{block}
+
+\begin{block}{Outlook}
+\begin{itemize}
+    \item there exists a $1.375$-approximation (\citeauthor*{Berman2001})
+    \item MIN-SBR for signed permutations is in $P$ (\citeauthor*{Hannenhalli1995})
+    \item MIN-SBR for signed permutations is in $P$ (\citeauthor*{Hannenhalli1995})
+\end{itemize}
+\end{block}
+
+\end{frame}
+
 \begin{frame}[allowframebreaks]{References}
 
+\nocite{*}
 \printbibliography

### 2020-07-07 slides.tex
update slides

@@
 \end{align*}
-and $d$ is minimal. \pause\par\padding
+and $d$ is minimal. \pause\par\spadding
 $d$ is called \b{reversal distance} between $\sigma$ and $\tau$.
-\end{definition}
+\end{definition}\pause\padding
+
+Observation: The reversal distance between $\sigma$ and $\tau$ is the same as the reversal distance between $\tau^{-1} \circ \sigma$ and $id$.
 
@@
 
-Observation: The reversal distance between $\sigma$ and $\tau$ is the same as the reversal distance between $\tau^{-1} \circ \sigma$ and $id$. \pause\padding
-
 \begin{definition}[MIN-SBR]
 Let $\pi = \tau^{-1} \circ \sigma \in S_n$. \par\pause
-\textit{Sorting by Reversals} is the problem of finding a sequence of reversals $\rho_1, \dots, \rho_d \in S_n$ such that
+\b{Sorting by Reversals} is the problem of finding a sequence of reversals $\rho_1, \dots, \rho_d \in S_n$ such that
 \begin{align*}
@@
 and $d$ is minimal.
-\end{definition}
-
-\end{frame}
-
-\begin{frame}
+\end{definition}\pause
 
@@
 
+\end{frame}
+
+% \begin{frame}
+
 % \begin{algorithm}[ratchet algorithm, \citeauthor*{Watterson19821}]
@@
 
-\end{frame}
+% \end{frame}
 
@@
 \begin{frame}
+
+A different perspective: $\pi = (0\ 1\ \r\mid\ 3\ 4\ \r\mid\ 2\ \r\mid\ 5)$\pause\spadding
 
@@
 
-\begin{proof}\renewcommand{\qedsymbol}{}
+\begin{proof}
 Let $\pi_t = \pi, \pi_0 = id$ and $\rho_1, \dots, \rho_t$ a shortest sequence of reversals from $\pi_t$ to $\pi_0$. \pause Then
@@
     \end{aligned}
-\end{align*}
+\end{align*}\pause
+Setting $i=t$, proves the theorem.
 \end{proof}
@@
 \begin{frame}
-
-\begin{proof}[Proof (cont.)]
-Let $i = t$. Then
-\begin{align*}
-         && d(\pi) - (b(\pi) - c(\pi)) &\geq 0 \\ \pause
-    \iff && d(\pi)                     &\geq b(\pi) - c(\pi).
-\end{align*}
-\end{proof}\pause
 
@@
 
-\begin{frame}
+% \begin{frame}
 
@@
 \begin{proof}[Proof (cont.)]
-To show: $k+l+m-v \geq \frac{c_2(\pi)}{2}$. \pause We know that
+Left to show: $k+l+m-v \geq \frac{c_2(\pi)}{2}$. \pause We know that
 \begin{enumerate}
@@
 \begin{itemize}
-    \item the number of alternating cycles in a breakpoint graph $G(\pi)$ are related to $d(\pi)$
+    \item the number of alternating cycles in a breakpoint graph $G(\pi)$ is related to $d(\pi)$
     \item a sorting sequence of reversals can be constructed from an alternating cycle decomposition of $G(\pi)$
@@
     \item there exists a $1.375$-approximation (\citeauthor*{Berman2001})
-    \item MIN-SBR for signed permutations is in $P$ (\citeauthor*{Hannenhalli1995})
     \item MIN-SBR for signed permutations is in $P$ (\citeauthor*{Hannenhalli1995})

### 2020-07-08 slides.tex
add citation
update slides

@@
 \begin{definition}
-Let $i \sim j$ if $|i - j| = 1$. \par\pause
+Let $i \sim j$ if $|i - j| = 1$. \par
 A pair of consecutive elements $\pi_i$ and $\pi_j$ is
@@
 \begin{itemize}
-    \item Every connected component $A$ arising from k different alternating cycles of $G(\pi)$, eventually reduces to a two cycle.\pause
+    \item Every connected component $A$ arising from $k$ different alternating cycles of $G(\pi)$, eventually reduces to $k$ $2$-cycles.\pause
     \item There exists an elimination sequence of $A$ with $k$ $2$-reversals and remaining $1$-reversals.\pause
@@
 \begin{enumerate}
-    \item $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ as $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles by $M$\pause;
+    \item $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ as $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles in $C$\pause;
     \item $v \leq \left\lfloor \frac{z}{2} \right\rfloor$ as every unoriented component representing a $2$-cycle represents at least one more alternating cycle (lemma \ref{lem:7})\pause; and
@@
 
-Run time: $O(n^4)$, can be improved to $O(n^2)$.\spadding
+Run time: $O(n^4)$, can be improved to $O(n^2)$ (\citeauthor*{Kaplan1997}).\spadding
 

### 2020-07-08 slides.tex
update slides

@@
 \def\r{\textcolor{red}}
+\def\g#1{{\usebeamercolor[fg]{block title example}{#1}}}
 
@@
 \end{itemize} \par\pause\spadding
-$b(\pi)$ denotes the number of breakpoints in $\pi \in S_n$.
+$\r{b(\pi)}$ denotes the number of breakpoints in $\pi \in S_n$.
 \end{definition}\pause\padding
 
-Observation: $b(\pi) = 0$ iff $\pi = id$ \pause and any reversal can at most eliminate two breakpoints. \pause\spadding
+Observation: $\r{b(\pi)} = 0$ iff $\pi = id$ \pause and any reversal can at most eliminate two breakpoints. \pause\spadding
 
 \begin{corollary}[lower bound, \citeauthor*{Kececioglu1995}]
-$d(\pi) \geq \left\lceil \frac{b(\pi)}{2} \right\rceil$ for all $\pi \in S_n$.
+$d(\pi) \geq \left\lceil \frac{\r{b(\pi)}}{2} \right\rceil$ for all $\pi \in S_n$.
 \end{corollary}
@@
 		\node [style=node] (0) at (-1.25, 0) {i};
+		\node [style=node] (1) at (1.25, 0) {i'};
+		\node [style=node] (2) at (3, 0) {j'};
+		\node [style=node] (3) at (-3, 0) {j};
+		\node (4) at (0, 0) {$\dots$};
+		\node (5) at (-4.25, 0) {$\dots$};
+		\node (6) at (4.25, 0) {$\dots$};
+	\end{pgfonlayer}
+	\begin{pgfonlayer}{edgelayer}
+		\draw [style=blue edge, bend right] (2) to (3);
+		\draw [style=blue edge, bend left] (0) to (1);
+		\draw [style=red edge] (3) to (0);
+		\draw [style=red edge] (1) to (2);
+	\end{pgfonlayer}
+\end{tikzpicture}
+\end{center}
+\caption{Unoriented $2$-cycle}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+	\begin{pgfonlayer}{nodelayer}
+		\node [style=node] (0) at (-1.25, 0) {i};
 		\node [style=node] (1) at (1.25, 0) {j'};
@@
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=node] (0) at (-1.25, 0) {i};
-		\node [style=node] (1) at (1.25, 0) {i'};
-		\node [style=node] (2) at (3, 0) {j'};
-		\node [style=node] (3) at (-3, 0) {j};
-		\node (4) at (0, 0) {$\dots$};
-		\node (5) at (-4.25, 0) {$\dots$};
-		\node (6) at (4.25, 0) {$\dots$};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw [style=blue edge, bend right] (2) to (3);
-		\draw [style=blue edge, bend left] (0) to (1);
-		\draw [style=red edge] (3) to (0);
-		\draw [style=red edge] (1) to (2);
-	\end{pgfonlayer}
-\end{tikzpicture}
-\end{center}
-\caption{Unoriented $2$-cycle}
-\end{figure}
-
 \end{frame}
@@
 
-Let $c(\pi)$ denote the maximum number of alternating cycles in any alternating cycle decomposition of $G(\pi)$. \pause\padding
+Let $\g{c(\pi)}$ denote the maximum number of alternating cycles in any alternating cycle decomposition of $G(\pi)$. \pause\padding
 
@@
 \begin{align*}
-    b(\pi) - b(\pi \circ \rho) + c(\pi \circ \rho) - c(\pi) \leq 1.
+    \r{b(\pi)} - \r{b(\pi \circ \rho)} + \g{c(\pi \circ \rho)} - \g{c(\pi)} \leq 1.
 \end{align*}
@@
 \begin{proof}
-To show: $b(\pi) - b(\pi \circ \rho) + c(\pi \circ \rho) - c(\pi) \leq 1$. \par
-We consider each case $b(\pi) - b(\pi \circ \rho) \in [-2, 2]$ separately.\pause
+To show: $\r{b(\pi)} - \r{b(\pi \circ \rho)} + \g{c(\pi \circ \rho)} - \g{c(\pi)} \leq 1$. \par
+We consider each case $\r{b(\pi)} - \r{b(\pi \circ \rho)} \in [-2, 2]$ separately.\pause
 \begin{enumerate}
@@
 \begin{align*}
-    d(\pi) \geq b(\pi) - c(\pi).
+    d(\pi) \geq \r{b(\pi)} - \g{c(\pi)}.
 \end{align*}
@@
     d(\pi_i) &= d(\pi_{i-1}) + 1 \\ \pause
-             \overset{(\ref{thm:1})}&{\geq} d(\pi_{i-1}) + b(\pi_i) - b(\pi_{i-1}) + c(\pi_{i-1}) - c(\pi_i) \\[0.25cm] \pause
+             \overset{(\ref{thm:1})}&{\geq} d(\pi_{i-1}) + \r{b(\pi_i)} - \r{b(\pi_{i-1})} + \g{c(\pi_{i-1})} - \g{c(\pi_i)} \\[0.25cm] \pause
     \iff & \begin{aligned}[t]
-    d(\pi_i) - (b(\pi_i) - c(\pi_i)) &\geq d(\pi_{i-1}) - (b(\pi_{i-1}) - c(\pi_{i-1})) \\ \pause
-                                     &\geq d(\pi_0) - (b(\pi_0) - c(\pi_0))\pause = 0
+    d(\pi_i) - (\r{b(\pi_i)} - \g{c(\pi_i)}) &\geq d(\pi_{i-1}) - (\r{b(\pi_{i-1})} - \g{c(\pi_{i-1})}) \\ \pause
+                                     &\geq d(\pi_0) - (\r{b(\pi_0)} - \g{c(\pi_0)})\pause = 0
     \end{aligned}
@@
 \label{thm:3}
-Let $\pi \in S_n$ and $C$ be a maximum alternating cycle decomposition of $G(\pi)$. Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in any such $C$. Then
+Let $\pi \in S_n$ and $C$ be a maximum alternating cycle decomposition of $G(\pi)$. Let $\g{c_2(\pi)}$ be the minimum number of alternating $2$-cycles in any such $C$. Then
 \begin{align*}
-    d(\pi) \geq \frac{2}{3} b(\pi) - \frac{1}{3} c_2(\pi).
+    d(\pi) \geq \frac{2}{3} \r{b(\pi)} - \frac{1}{3} \g{c_2(\pi)}.
 \end{align*}
@@
 \begin{block}{$\frac{3}{2}$-approximation}
-By theorem \ref{thm:3}, an algorithm that finds a sorting sequence of reversals of at most length $b(\pi) - \frac{1}{2} c_2(\pi)$ achieves an approximation bound of $\frac{3}{2}$. \pause\spadding
+By theorem \ref{thm:3}, an algorithm that finds a sorting sequence of reversals of at most length $\r{b(\pi)} - \frac{1}{2} \g{c_2(\pi)}$ achieves an approximation bound of $\frac{3}{2}$.
+\pause\spadding
 
@@
 
+\begin{figure}
 \begin{center}
-\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
+\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 \end{tikzpicture}
-\end{center}\pause
+\end{center}
+\caption{Unoriented $2$-cycle}
+\end{figure}
 
-\begin{proof}
-Let $u$ and $v$ represent the blue edges of an unoriented $2$-cycle of $C$. \par\pause
-Then its two constituting blue edges do not interleave. Hence, $R(C)$ does not connect $u$ and $v$ with an edge. \par\pause
-If its constituting blue edges do not interleave with blue edges of any other alternating cycle of $C$, then $u$ and $v$ would be isolated blue vertices. \b{$\lightning$}
-\end{proof}
+% \begin{proof}
+% Let $u$ and $v$ represent the blue edges of an unoriented $2$-cycle of $C$. \par\pause
+% Then its two constituting blue edges do not interleave. Hence, $R(C)$ does not connect $u$ and $v$ with an edge. \par\pause
+% If its constituting blue edges do not interleave with blue edges of any other alternating cycle of $C$, then $u$ and $v$ would be isolated blue vertices. \b{$\lightning$}
+% \end{proof}
 
@@
 \begin{align*}
-    d(\pi) \leq b(\pi) - |C| + u(C)
+    d(\pi) \leq \r{b(\pi)} - |C| + u(C)
 \end{align*}
@@
 \begin{align*}
-    d(\pi) \leq b(\pi) - \frac{1}{2} c_2(\pi).
+    d(\pi) \leq \r{b(\pi)} - \frac{1}{2} \g{c_2(\pi)}.
 \end{align*}
@@
 \begin{align*}
-    d(\pi) &\leq b(\pi) - k - l - u - m + u + v \\
-           &= b(\pi) - k - l - m + v
-\end{align*}
+    d(\pi) &\leq \r{b(\pi)} - k - l - u - m + u + v \\
+           &= \r{b(\pi)} - k - l - m + v
+\end{align*}\pause
+Left to show: $-k-l-m+v \leq - \frac{1}{2} \g{c_2(\pi)}$.
 \end{proof}
@@
 \begin{proof}[Proof (cont.)]
-Left to show: $k+l+m-v \geq \frac{c_2(\pi)}{2}$. \pause We know that
+Left to show: $k+l+m-v \geq \frac{1}{2} \g{c_2(\pi)}$. \pause We know that
 \begin{enumerate}
@@
             &= \left\lceil \frac{y}{2} \right\rceil + \left\lceil \frac{z}{2} \right\rceil \\ \pause
-            &\geq \frac{c_2(\pi)}{2}                                                                    & (3)
+            &\geq \frac{1}{2} \g{c_2(\pi)}                                                              & (3)
 \end{align*}

### 2020-07-08 slides.tex
update slides

@@
 \tikzstyle{node}=[fill=white, draw=black, shape=circle, minimum size=0.75cm]
-\tikzstyle{blue node}=[fill=white, draw=blue, shape=circle]
-\tikzstyle{red node}=[fill=white, draw=red, shape=circle]
-\tikzstyle{blue edge}=[-, draw=blue]
-\tikzstyle{red edge}=[-, draw=red]
+\tikzstyle{blue node}=[fill=white, draw=blue, shape=circle, line width=0.25mm]
+\tikzstyle{red node}=[fill=white, draw=red, shape=circle, line width=0.25mm]
+\tikzstyle{blue edge}=[-, draw=blue, line width=0.25mm]
+\tikzstyle{red edge}=[-, draw=red, line width=0.25mm]
+\tikzstyle{dashed edge}=[-, dashed, line width=0.25mm]
 \pgfdeclarelayer{nodelayer}
@@
 
-An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges. \pause\spadding
+An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges.
+\end{definition}
 
-We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$.
-\end{definition}
+% We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$.
 
@@
 
-\begin{frame}
+% \begin{frame}
 
-\begin{lemma}[\citeauthor*{Christie1998}]
-\label{lem:2}
-An oriented alternating cycle is a $2$-cycle iff the corresponding reversal acting on both red edges is a $2$-reversal.
-\end{lemma}\spadding
+% \begin{lemma}[\citeauthor*{Christie1998}]
+% \label{lem:2}
+% An oriented alternating cycle is a $2$-cycle iff the corresponding reversal acting on both red edges is a $2$-reversal.
+% \end{lemma}\spadding
 
-Let $i \sim i'$ and $j \sim j'$.
+% Let $i \sim i'$ and $j \sim j'$.
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=node] (0) at (-1.25, 0) {i};
-		\node [style=node] (1) at (1.25, 0) {i'};
-		\node [style=node] (2) at (3, 0) {j'};
-		\node [style=node] (3) at (-3, 0) {j};
-		\node (4) at (0, 0) {$\dots$};
-		\node (5) at (-4.25, 0) {$\dots$};
-		\node (6) at (4.25, 0) {$\dots$};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw [style=blue edge, bend right] (2) to (3);
-		\draw [style=blue edge, bend left] (0) to (1);
-		\draw [style=red edge] (3) to (0);
-		\draw [style=red edge] (1) to (2);
-	\end{pgfonlayer}
-\end{tikzpicture}
-\end{center}
-\caption{Unoriented $2$-cycle}
-\end{figure}
+% \begin{figure}
+% \begin{center}
+% \begin{tikzpicture}
+% 	\begin{pgfonlayer}{nodelayer}
+% 		\node [style=node] (0) at (-1.25, 0) {i};
+% 		\node [style=node] (1) at (1.25, 0) {i'};
+% 		\node [style=node] (2) at (3, 0) {j'};
+% 		\node [style=node] (3) at (-3, 0) {j};
+% 		\node (4) at (0, 0) {$\dots$};
+% 		\node (5) at (-4.25, 0) {$\dots$};
+% 		\node (6) at (4.25, 0) {$\dots$};
+% 	\end{pgfonlayer}
+% 	\begin{pgfonlayer}{edgelayer}
+% 		\draw [style=blue edge, bend right] (2) to (3);
+% 		\draw [style=blue edge, bend left] (0) to (1);
+% 		\draw [style=red edge] (3) to (0);
+% 		\draw [style=red edge] (1) to (2);
+% 	\end{pgfonlayer}
+% \end{tikzpicture}
+% \end{center}
+% \caption{Unoriented $2$-cycle}
+% \end{figure}
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-	\begin{pgfonlayer}{nodelayer}
-		\node [style=node] (0) at (-1.25, 0) {i};
-		\node [style=node] (1) at (1.25, 0) {j'};
-		\node [style=node] (2) at (3, 0) {i'};
-		\node [style=node] (3) at (-3, 0) {j};
-		\node (4) at (0, 0) {$\dots$};
-		\node (5) at (-4.25, 0) {$\dots$};
-		\node (6) at (4.25, 0) {$\dots$};
-	\end{pgfonlayer}
-	\begin{pgfonlayer}{edgelayer}
-		\draw [style=blue edge, bend right] (3) to (1);
-		\draw [style=blue edge, bend left] (0) to (2);
-		\draw [style=red edge] (3) to (0);
-		\draw [style=red edge] (1) to (2);
-	\end{pgfonlayer}
-\end{tikzpicture}
-\end{center}
-\caption{Oriented $2$-cycle}
-\end{figure}
+% \begin{figure}
+% \begin{center}
+% \begin{tikzpicture}
+% 	\begin{pgfonlayer}{nodelayer}
+% 		\node [style=node] (0) at (-1.25, 0) {i};
+% 		\node [style=node] (1) at (1.25, 0) {j'};
+% 		\node [style=node] (2) at (3, 0) {i'};
+% 		\node [style=node] (3) at (-3, 0) {j};
+% 		\node (4) at (0, 0) {$\dots$};
+% 		\node (5) at (-4.25, 0) {$\dots$};
+% 		\node (6) at (4.25, 0) {$\dots$};
+% 	\end{pgfonlayer}
+% 	\begin{pgfonlayer}{edgelayer}
+% 		\draw [style=blue edge, bend right] (3) to (1);
+% 		\draw [style=blue edge, bend left] (0) to (2);
+% 		\draw [style=red edge] (3) to (0);
+% 		\draw [style=red edge] (1) to (2);
+% 	\end{pgfonlayer}
+% \end{tikzpicture}
+% \end{center}
+% \caption{Oriented $2$-cycle}
+% \end{figure}
 
-\end{frame}
+% \end{frame}
 
@@
             		\draw [style=red edge] (1) to (2);
-            		\draw [dashed, bend left] (0) to (2);
+            		\draw [style=dashed edge, bend left] (0) to (2);
             	\end{pgfonlayer}
@@
             		\draw [style=red edge] (1) to (2);
-            		\draw [dashed, bend left] (1) to (2);
+            		\draw [style=dashed edge, bend left] (1) to (2);
             	\end{pgfonlayer}
@@
 \begin{corollary}[\citeauthor*{Christie1998}]
-$R(C)$ consists of $n+1$ isolated blue vertices.
+$R(\emptyset)$ consists of $n+1$ isolated blue vertices.
 \end{corollary}

### 2020-07-09 slides.tex
update slides

@@
 
-An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges.
+An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges. \pause\spadding
+
+We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$.
 \end{definition}
-
-% We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$.
 
@@
 \begin{lemma}[\citeauthor*{Christie1998}]
+\label{lem:3}
+All vertices arising from the same alternating cycle in $C$ are in the same connected component of $R(C)$.
+\end{lemma}\pause
+
+\begin{lemma}[\citeauthor*{Christie1998}]
 \label{lem:7}
@@
 
-\begin{lemma}[\citeauthor*{Christie1998}]
-\label{lem:3}
-All vertices arising from the same alternating cycle in $C$ are in the same connected component of $R(C)$.
-\end{lemma}\pause
-
 \begin{definition}
@@
 \begin{exampleblock}{Example (elimination sequence)}
-Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$\onslide<2->{ and $u_1 = \{1,2\}$}\onslide<3->{, $u_2 = \{2,3\}$.}\spadding
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$\spadding
 
-\only<1>{\begin{center}\begin{tikzpicture}
+\begin{center}\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}\end{center}}\spadding
-\only<1>{\begin{center}\begin{tikzpicture}
+\end{tikzpicture}\end{center}\spadding
+\begin{center}\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}\end{center}}
-\only<2>{\begin{center}\begin{tikzpicture}
+\end{tikzpicture}\end{center}
+\end{exampleblock}
+
+\end{frame}
+
+\begin{frame}
+
+\begin{exampleblock}{Example (elimination sequence)}
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$ and $u_1 = \{1,2\}$\spadding
+
+\begin{center}\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}\end{center}}\spadding
-\only<2>{\begin{center}\begin{tikzpicture}
+\end{tikzpicture}\end{center}\spadding
+\begin{center}\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}\end{center}}
-\only<3>{\begin{center}\begin{tikzpicture}
+\end{tikzpicture}\end{center}
+\end{exampleblock}
+
+\end{frame}
+
+\begin{frame}
+
+\begin{exampleblock}{Example (elimination sequence)}
+Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$ and $u_1 = \{1,2\}$, $u_2 = \{2,3\}$.\spadding
+
+\begin{center}\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}\end{center}}\spadding
-\only<3>{\begin{center}\begin{tikzpicture}
+\end{tikzpicture}\end{center}\spadding
+\begin{center}\begin{tikzpicture}
 	\begin{pgfonlayer}{nodelayer}
@@
 	\end{pgfonlayer}
-\end{tikzpicture}\end{center}}
+\end{tikzpicture}\end{center}
 \end{exampleblock}
@@
     \item $v \leq \left\lfloor \frac{z}{2} \right\rfloor$ as every unoriented component representing a $2$-cycle represents at least one more alternating cycle (lemma \ref{lem:7})\pause; and
-    \item $|M| = y+z \geq c_2(\pi)$.
+    \item $|M| = y+z \geq \g{c_2(\pi)}$.
 \end{enumerate}\pause

### 2020-07-11 paper.tex
update examples
paper

@@
-\documentclass{article}
+\documentclass[11pt,DIV=11]{scrartcl}
 \usepackage[utf8]{inputenc}
 \usepackage[english]{babel}
+\usepackage[ruled]{algorithm2e}
+\usepackage{aligned-overset}
 \usepackage{alltt}
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{csquotes}
+\usepackage{enumerate}
 \usepackage{multicol}
+\usepackage{stmaryrd}
+\usepackage{subcaption}
+\usepackage{xcolor}
 
-\def\code#1{\texttt{#1}}
+% \setlength\parskip{0.2em}
+\binoppenalty=32767
+\relpenalty=32767
 
-\usepackage{biblatex}
+\def\code#1{\texttt{\frenchspacing#1}}
+\def\padding{\vspace{3em}}
+
+\usepackage{tikz}
+\tikzstyle{node}=[fill=white, draw=black, shape=circle, minimum size=2em]
+\tikzstyle{blue node}=[fill=white, draw=blue, shape=circle]
+\tikzstyle{red node}=[fill=white, draw=red, shape=circle]
+\tikzstyle{blue edge}=[-, draw=blue]
+\tikzstyle{red edge}=[-, draw=red]
+\tikzstyle{dashed edge}=[-, dashed]
+\pgfdeclarelayer{nodelayer}
+\pgfdeclarelayer{edgelayer}
+\pgfsetlayers{nodelayer,edgelayer}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}{Example}[theorem]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+
+\usepackage[sorting=ynt,style=alphabetic]{biblatex}
 \addbibresource{sources.bib}
-
-\evensidemargin=0in
-\oddsidemargin=0in
-\textwidth=6.5in
-\textheight=9.0in
 
@@
 
+\begin{abstract}
+    A reversal is a transformation, reversing the order of elements of a permutation.
+Sorting by Reversals is the problem of finding the shortest sequence of reversals transforming a permutation to the identity.
+The length of such a sequence is said to be the reversal distance of that permutation.
+In this paper we introduce fundamental concepts used in the literature on this topic.
+In particular, the fundamental relationship between the reversal distance and the number of alternating cycles in a cycle decomposition of a suitably-defined bi-colored graph.
+Then a $3/2$-approximation is discussed in detail.
+We also reference further related research.
+\end{abstract}
+
+\section{Introduction}
+
+In computational biology a fundamental problem is to identify the evolutionary similarity between genome sequences.
+By far the most common mutation on genomes is the inversion.
+A genome can be represented by a permutation.
+An inversion then simply is the reversal of the substring of a permutation \cite{Kececioglu1995}.
+
+We refer by $\langle S_n, \circ \rangle$ to the symmetric group with $S_n = \{(\pi_1\ \dots\ \pi_n) \mid \{\pi_1, \dots, \pi_n\} = [1,n]\}$.
+Let $\pi \in S_n$ be a permutation and $\pi_i$ denote $\pi(i)$.
+The identity permutation $id = (1\ \dots\ n)$ is the permutation mapping every input to itself.
+A \textit{reversal} $\rho(i,j)$ inverses the substring from $\pi_i$ to $\pi_j$ of a permutation.
+It is represented by the permutation
+\begin{align*}
+    (1\ \cdots\ i-1\ j\ j-1\ \cdots\ i+1\ i\ j+1\ \cdots\ n).
+\end{align*}
+Composing $\pi$ with $\rho(i,j)$ yields $(\pi_1\ \cdots\ \pi_{i-1}\ \pi_j\ \pi_{j-1}\ \dots\ \pi_{i+1}\ \pi_i\ \pi_{j+1}\ \cdots\ \pi_n)$ where the elements $\pi_i, \dots, \pi_j$ have been reversed.
+
+Motivated by the outlined application in computational biology, we define the following problem.
+
+\begin{definition}
+Given two permutations $\sigma$ and $\tau$, the \textit{reversal distance problem} is the problem of finding the shortest sequence of reversals transforming $\sigma$ into $\tau$.
+More formally, we want to find reversals $\rho_1, \dots, \rho_d$ such that
+\begin{align*}
+    \sigma \circ \rho_1 \circ \cdots \circ \rho_d = \tau
+\end{align*}
+and $d$ is minimal.
+The number of necessary reversals $d$ is called the \textit{reversal distance} between $\sigma$ and $\tau$.
+\end{definition}
+
+We now observe that the very same shortest sequence of reversals transforming $\sigma$ into $\tau$ is also the shortest sequence of reversals transforming $\tau^{-1} \circ \sigma$ into $id$.
+This simply follows from left-composing $\tau^{-1}$ with both sides of the equation.
+
+\begin{definition}
+\textit{Sorting by Reversals (MIN-SBR)} then is the problem of determining the reversal distance between some permutation $\pi$ and $id$.
+The reversal distance is denoted by $d(\pi)$.
+We say that a sequence of reversals transforming $\pi$ into $id$, \textit{sorts} $\pi$.
+\end{definition}
+
+Perhaps the most natural algorithm for sorting by reversals is the \textit{ratchet algorithm} proposed by \citeauthor*{Watterson19821} \cite{Watterson19821}.
+The algorithm sequentially brings all elements from $1$ to $n$ into the correct position.
+More formally, in every step $i$, if $\pi_i \neq i$, we apply the reversal $\rho(i, \pi_i^{-1})$ ($\pi_i^{-1}$ refers to the position of element $i$ in the permutation $\pi$).
+Once step $n-1$ is completed, element $n$ must be in position $n$.
+Therefore, this algorithm sorts any permutation in at most $n-1$ steps.
+The approximation of this algorithm, however, can be arbitrarily poor.
+Consider the permutation $(1\ n\ n-1\ \dots\ 2)$.
+The ratchet algorithm sorts this permutation in $n-1$ reversals.
+But the permutation can also be sorted by the single reversal $\rho(2, n)$ \cite{Kececioglu1995}.
+
+MIN-SBR has been proven NP-hard by \citeauthor*{Caprara1997} \cite{Caprara1997} and MAX-SNP hard by \citeauthor*{Berman1999} excluding the existance of a polynomial time approximation scheme for this problem \cite{Berman1999}.
+The problem of sorting signed permutations by reversals, that is permutations where each element may be either positive or negative, was shown to be solvable in polynomial time by \citeauthor*{Hannenhalli1995} \cite{Hannenhalli1995}.
+
+In this paper we give an overview of approximation algorithms for MIN-SBR.
+In section \ref{sec:breakpoints} we first introduce the notion of breakpoints introduced by \citeauthor*{Kececioglu1995} that allow us to formulate a first lower bound for the reversal distance of $\pi$ and lead to a $2$-approximation \cite{Kececioglu1995}.
+In section \ref{sec:breakpoint_graph} we define a suitable bi-colored graph first introduced by \citeauthor*{Bafna1996} to find a tighter lower bound \cite{Bafna1996}.
+In this section, we make a fundamental observation relating the reversal distance to the number of alternating cycles in a cycle decomposition of such a graph.
+Lastly, we consider a $3/2$-approximation described by \citeauthor{Christie1998} \cite{Christie1998} (Section \ref{sec:32_approximation}).
+A $1.375$-approximation was later obtained by \citeauthor*{Berman2001}, but is not considered here \cite{Berman2001}.
+
+\section{Breakpoints}
+\label{sec:breakpoints}
+
+Now we consider the relationship between two consecutive elements of a permutation.
+
+\begin{definition}
+Let $i \sim j$ if $|i - j| = 1$.
+A pair of consecutive elements $\pi_i$ and $\pi_j$ forms an \textit{adjacency} if $\pi_i \sim \pi_j$ and a \textit{breakpoint} if $\pi_i \not\sim \pi_j$.
+We denote by $b(\pi)$ the number of breakpoints in the permutation $\pi$.
+Figure \ref{fig:breakpoints} shows an example for the breakpoints of a permutation.
+\end{definition}
+
+\begin{figure}
+    \begin{align*}
+        \pi = (1\ \mid\ 3\ 4\ \mid\ 2)
+    \end{align*}
+    \caption{Breakpoints (breakpoints marked by vertical bars)}
+    \label{fig:breakpoints}
+\end{figure}
+
+We observe that the number of breakpoints of the identity permutation $b(id)$ is $0$.
+Based on this observation our goal in the following sections is to reduce the number of breakpoints of a given permutation $\pi$ to $0$ as quickly as possible by applying reversals.
+We also observe that with our current definition of permutations, the identity permutation is in fact not the only permutation without any breakpoints.
+Consider the permutation $(n\ n+1\ \dots\ 2\ 1)$ (the inversed identity permutation) which also does not have any breakpoints.
+
+In an effort to make the identity permutation the only permutation without any breakpoints, we extend our definition of permutations (and the symmetric group) by expanding permutations with one fixed initial $0$ and one fixed trailing $n+1$.
+In the following we only consider such expanded permutations $\pi = (0\ \pi_1\ \dots\ \pi_n\ n+1) \in S_n$ where $\pi_0 = 0$ and $\pi_{n+1} = n+1$.
+It is easy to see that for an expanded permutation $\pi$, $b(\pi) = 0$ if and only if $\pi = id$.
+
+We call a reversal a \textit{$k$-reversal} if it removes $k$ breakpoints. It is immediate that $k \in [-2,2]$ as a reversal can at most add or eliminate $2$ breakpoints.
+
+\begin{corollary}
+Using the previous two observations, \citeauthor*{Kececioglu1995} found a first fundamental lower bound to the reversal distance of a permutation $\pi$ \cite{Kececioglu1995}.
+\begin{align*}
+    d(\pi) \geq \left\lceil \frac{b(\pi)}{2} \right\rceil.
+\end{align*}
+\end{corollary}
+
+Based on this first lower bound, \citeauthor*{Kececioglu1995} proposed a simple $2$-approximation running in $O(n^2)$ \cite{Kececioglu1995}.
+
+\section{Breakpoint graph}
+\label{sec:breakpoint_graph}
+
+We now try to find a tighter lower bound by introducing the breakpoint graph of a permutation.
+
+\begin{definition}
+The \textit{breakpoint graph} $G(\pi)$ has as vertices the elements of the permutation $\pi$, $[0, n+1]$.
+We connect two vertices $\pi_i$ and $\pi_j$ with a \textit{red} edge if they form a breakpoint in $\pi$.
+And we connect two vertices $\pi_i$ and $\pi_j$ representing non-consecutive elements in $\pi$ (that is $|\pi_i^{-1} - \pi_j^{-1}| \neq 1$) with a \textit{blue} edge if $\pi_i \sim \pi_j$.
+\end{definition}
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}
+    	\begin{pgfonlayer}{nodelayer}
+    		\node [style=node] (0) at (-3, 0) {0};
+    		\node [style=node] (1) at (-1.5, 0) {1};
+    		\node [style=node] (2) at (0, 0) {3};
+    		\node [style=node] (3) at (1.5, 0) {4};
+    		\node [style=node] (4) at (3, 0) {2};
+    		\node [style=node] (5) at (4.5, 0) {5};
+    	\end{pgfonlayer}
+    	\begin{pgfonlayer}{edgelayer}
+    		\draw [style=red edge] (1) to (2);
+    		\draw [style=red edge] (3) to (4);
+    		\draw [style=red edge] (4) to (5);
+    		\draw [style=blue edge, bend right] (1) to (4);
+    		\draw [style=blue edge, bend left] (2) to (4);
+    		\draw [style=blue edge, bend left] (3) to (5);
+    	\end{pgfonlayer}
+    \end{tikzpicture}
+    \caption{Breakpoint graph}
+    \label{fig:breakpoint_graph}
+\end{figure}
+
+The red edges in the breakpoint graph identify the elements of $\pi$ that have to be split apart and the blue edges identify the elements of $\pi$ that have to be brought together (the missing adjacencies) in the process of sorting the permutation $\pi$.
+Figure \ref{fig:breakpoint_graph} shows the breakpoint graph of the permutation $(0\ 1\ 3\ 4\ 2\ 5)$.
+
+An \textit{alternating cycle decomposition} of $G(\pi)$ is a collection of edge-disjoint alternating cycles, such that every edge of $G(\pi)$ is contained in exactly one cycle of the collection.
+We observe that for every breakpoint, a vertice in $G(\pi)$ must have a missing adjacency.
+And conversely, for every missing adjacency a vertex in $G(\pi)$ must have a breakpoint.
+Therefore, each vertex in our breakpoint graph has an equal number of incident red and blue edges.
+
+\begin{corollary}
+Hence, $G(\pi)$ can be decomposed into edge-disjoint alternating cycles.
+\end{corollary}
+
+The main result we prove in this section, relates the maximum number of alternating cycles of any cycle decomposition of $G(\pi)$ to the reversal distance of $\pi$.
+
+First, we need to define some terminology.
+An alternating cycle in $G(\pi)$ is a \textit{$k$-cycle} if it has $k$ constituting red and $k$ constituting blue edges.
+We say, a reversal \textit{acts on} two red edges of $G(\pi)$ if those two red edges represent the breakpoints that are split apart by the reversal.
+Now, an alternating cycle in $G(\pi)$ is \textit{oriented} if there is a $1$- or a $2$-reversal acting on two constituting red edges.
+An alternating cycle of $G(\pi)$ is \textit{unoriented} if it is not oriented.
+
+To give a more visual means of deducing whether an alternating cycle of $G(\pi)$ is oriented, we assign to each red edge $(\pi_i, \pi_{i+1})$ an orientation from $\pi_i$ to $\pi_{i+1}$, orienting red edges of $G(\pi)$ from the endpoint which appears first in $\pi$ to the endpoint which appears second.
+Then, we call an alternating cycle of $G(\pi)$ \textit{directed} with respect to $\pi$ if it is possible to walk along the whole cycle traversing each red edge in the direction of its orientation.
+An alternating cycle of $G(\pi)$ is called \textit{undirected} if it is not directed.
+For example, in figure \ref{fig:breakpoint_graph} the alternating cycle $(\{1,3\},\{2,3\},\{2,5\},\{4,5\},\{2,4\},\{1,2\})$ is directed with respect to $(0\ 1\ 3\ 4\ 2\ 5)$, whereas the alternating cycle $(\{1,3\},\{2,3\},\{2,4\},\{4,5\},\{2,5\},\{1,2\})$ is undirected with respect to $(0\ 1\ 3\ 4\ 2\ 5)$.
+\citeauthor*{Caprara1997} argues that an alternating cycle of $G(\pi)$ is oriented if and only if it is undirected \cite{Caprara1997}.
+
+\begin{lemma}
+\label{lem:1}
+An oriented alternating cycle is a $2$-cycle if and only if the corresponding reversal acting on both red edges is a $2$-reversal.
+\end{lemma}
+
+\begin{proof}
+Let $i \sim i'$ and $j \sim j'$.
+We observe that the $2$-cycles shown in figure \ref{fig:2cycles} are in fact the only types of $2$-cycles (with the exception of the symmetric case of an oriented $2$-cycle).
+Given an oriented $2$-cycle, we know that the corresponding reversal acting on its two constituting red edges brings $j'$ next to $j$ and $i$ next to $i'$, removing two breakpoints.
+
+On the other hand, given a $2$-reversal $\rho$, we know that $\rho$ has to act on two red edges of $G(\pi)$.
+Identify with $k$ and $l$ the vertices incident to the first red edge and with $m$ and $n$ the vertices incident to the second red edge.
+Now, because $\rho$ is removes both red edges from the breakpoint graph, we know that either $k \sim m$ and $l \sim n$ or $k \sim n$ and $l \sim m$.
+Thus, forming a $2$-cycle in $G(\pi)$.
+\end{proof}
+
+\begin{figure}
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+        	\begin{pgfonlayer}{nodelayer}
+        		\node [style=node] (0) at (-1.25, 0) {i};
+        		\node [style=node] (1) at (1.25, 0) {j'};
+        		\node [style=node] (2) at (3, 0) {i'};
+        		\node [style=node] (3) at (-3, 0) {j};
+        		\node (4) at (0, 0) {$\dots$};
+        		\node (5) at (-4.25, 0) {$\dots$};
+        		\node (6) at (4.25, 0) {$\dots$};
+        	\end{pgfonlayer}
+        	\begin{pgfonlayer}{edgelayer}
+        		\draw [style=blue edge, bend right] (3) to (1);
+        		\draw [style=blue edge, bend left] (0) to (2);
+        		\draw [style=red edge] (3) to (0);
+        		\draw [style=red edge] (1) to (2);
+        	\end{pgfonlayer}
+        \end{tikzpicture}
+        \caption{Oriented $2$-cycle}
+        \label{fig:2reversal}
+    \end{subfigure}
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+        	\begin{pgfonlayer}{nodelayer}
+        		\node [style=node] (0) at (-1.25, 0) {i};
+        		\node [style=node] (1) at (1.25, 0) {i'};
+        		\node [style=node] (2) at (3, 0) {j'};
+        		\node [style=node] (3) at (-3, 0) {j};
+        		\node (4) at (0, 0) {$\dots$};
+        		\node (5) at (-4.25, 0) {$\dots$};
+        		\node (6) at (4.25, 0) {$\dots$};
+        	\end{pgfonlayer}
+        	\begin{pgfonlayer}{edgelayer}
+        		\draw [style=blue edge, bend right] (2) to (3);
+        		\draw [style=blue edge, bend left] (0) to (1);
+        		\draw [style=red edge] (3) to (0);
+        		\draw [style=red edge] (1) to (2);
+        	\end{pgfonlayer}
+        \end{tikzpicture}
+        \caption{Unoriented $2$-cycle}
+    \end{subfigure}
+    \caption{$2$-cycles}
+    \label{fig:2cycles}
+\end{figure}
+
+A similar lemma is proven by \citeauthor*{Christie1998} which we use as a corollary of this lemma in section \ref{sec:reversal_graph} \cite{Christie1998}.\padding
+
+Let $c(\pi)$ denote the maximum number of alternating cycles in any alternating cycle decomposition of $G(\pi)$.
+\citeauthor*{Bafna1996} prove a fundamental theorem relating $b(\pi)$ with $c(\pi)$ \cite{Bafna1996}.
+
+\begin{theorem}
+\label{thm:1}
+Let $\pi, \rho \in S_n$ and $\rho$ be a reversal. Then
+\begin{align*}
+    b(\pi) - b(\pi \circ \rho) + c(\pi \circ \rho) - c(\pi) \leq 1.
+\end{align*}
+\end{theorem}
+
+\begin{figure}
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+        	\begin{pgfonlayer}{nodelayer}
+        		\node [style=node] (0) at (-1.25, 0) {i};
+        		\node [style=node] (1) at (1.25, 0) {j'};
+        		\node [style=node] (2) at (3, 0) {k};
+        		\node [style=node] (3) at (-3, 0) {j};
+        		\node (4) at (0, 0) {$\dots$};
+        		\node (5) at (-4.25, 0) {$\dots$};
+        		\node (6) at (4.25, 0) {$\dots$};
+        	\end{pgfonlayer}
+        	\begin{pgfonlayer}{edgelayer}
+        		\draw [style=blue edge, bend right] (3) to (1);
+        		\draw [style=red edge] (3) to (0);
+        		\draw [style=red edge] (1) to (2);
+        		\draw [style=dashed edge, bend left] (0) to (2);
+        	\end{pgfonlayer}
+        \end{tikzpicture}
+        \caption{before reversal}
+    \end{subfigure}
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+        	\begin{pgfonlayer}{nodelayer}
+        		\node [style=node] (0) at (-1.25, 0) {j'};
+        		\node [style=node] (1) at (1.25, 0) {i};
+        		\node [style=node] (2) at (3, 0) {k};
+        		\node [style=node] (3) at (-3, 0) {j};
+        		\node (4) at (0, 0) {$\dots$};
+        		\node (5) at (-4.25, 0) {$\dots$};
+        		\node (6) at (4.25, 0) {$\dots$};
+        	\end{pgfonlayer}
+        	\begin{pgfonlayer}{edgelayer}
+        		\draw [style=red edge] (1) to (2);
+        		\draw [style=dashed edge, bend left] (1) to (2);
+        	\end{pgfonlayer}
+        \end{tikzpicture}
+        \caption{after reversal}
+    \end{subfigure}
+    \caption{Case (i)}
+    \label{fig:case_i}
+\end{figure}
+
+\begin{figure}
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+        	\begin{pgfonlayer}{nodelayer}
+        		\node [style=node] (0) at (-3, 0) {i};
+        		\node [style=node] (1) at (-1.25, 0) {j};
+        		\node [style=node] (2) at (1.25, 0) {k};
+        		\node [style=node] (3) at (3, 0) {l};
+        		\node (4) at (0, 0) {$\dots$};
+        		\node (5) at (-4.25, 0) {$\dots$};
+        		\node (6) at (4.25, 0) {$\dots$};
+        	\end{pgfonlayer}
+        	\begin{pgfonlayer}{edgelayer}
+        		\draw [style=blue edge, bend right] (0) to (2);
+        		\draw [style=blue edge, bend right] (1) to (3);
+        		\draw [style=red edge] (2) to (3);
+        		\draw [style=dashed edge, bend left] (0) to (1);
+        	\end{pgfonlayer}
+        \end{tikzpicture}
+        \caption{before reversal}
+    \end{subfigure}
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
+        	\begin{pgfonlayer}{nodelayer}
+        		\node [style=node] (0) at (-3, 0) {i};
+        		\node [style=node] (1) at (-1.25, 0) {k};
+        		\node [style=node] (2) at (1.25, 0) {j};
+        		\node [style=node] (3) at (3, 0) {l};
+        		\node (4) at (0, 0) {$\dots$};
+        		\node (5) at (-4.25, 0) {$\dots$};
+        		\node (6) at (4.25, 0) {$\dots$};
+        	\end{pgfonlayer}
+        	\begin{pgfonlayer}{edgelayer}
+        		\draw [style=blue edge, bend right] (0) to (2);
+        		\draw [style=dashed edge, bend left] (0) to (2);
+        	\end{pgfonlayer}
+        \end{tikzpicture}
+        \caption{after reversal}
+    \end{subfigure}
+    \caption{Case (ii)}
+    \label{fig:case_ii}
+\end{figure}
+
+\begin{proof}
+We consider each case $b(\pi) - b(\pi \circ \rho) \in [-2,2]$ separately.
+\begin{enumerate}
+    \item Let $b(\pi) - b(\pi \circ \rho) = 2$.
+Lemma \ref{lem:1} implies that a $2$-reversal $\rho$ removes a $2$-cycle from every alternating cycle decomposition of $G(\pi)$.
+Therefore, $c(\pi \circ \rho) - c(\pi) \leq -1$.
+    \item Let $b(\pi) - b(\pi \circ \rho) = 1$. A $1$-reversal either creates one new adjacency (i) or creates two new adjacencies and simultaneously destroys one adjacency (ii).
+        \begin{enumerate}[(i)]
+            \item Let $j \sim j'$.
+Then figure \ref{fig:case_i} illustrates that every alternating cycle in $G(\pi \circ \rho)$ containing the edge $\{i,k\}$ induces an alternating cycle in $G(\pi)$ containing the edges $\{j,i\}$, $\{j,j'\}$, and $\{j',k\}$.
+Therefore, every alternating cycle decomposition of $G(\pi \circ \rho)$ into $c$ cycles induces an alternating cycle decomposition of $G(\pi)$ into $c$ cycles.
+            \item Let $i \sim j$, $i \sim k$ and $j \sim l$.
+Then figure \ref{fig:case_ii} illustrates that every alternating cycle in $G(\pi \circ \rho)$ containing the edge $\{i,j\}$ induces an alternating cycle in $G(\pi)$ containing the edges $\{i,k\}$, $\{k,l\}$, and $\{j,l\}$.
+Therefore, every alternating cycle decomposition of $G(\pi \circ \rho)$ into $c$ cycles induces an alternating cycle decomposition of $G(\pi)$ into $c$ cycles.
+        \end{enumerate}
+    Therefore, for both subcases, $c(\pi \circ \rho) - c(\pi) \leq 0$ holds.
+\end{enumerate}
+Proof for other cases similar.
+\end{proof}
+
+Using \ref{thm:1}, \citeauthor*{Bafna1996} give a lower bound of the reversal distance of a permutation $\pi$ in terms of $b(\pi)$ and $c(\pi)$ \cite{Bafna1996}.
+
+\begin{theorem}
+\label{thm:2}
+Let $\pi \in S_n$. Then
+\begin{align*}
+    d(\pi) \geq b(\pi) - c(\pi).
+\end{align*}
+\end{theorem}
+
+\begin{proof}
+Let $\pi_t = \pi, \pi_0 = id$ and $\rho_1, \dots, \rho_t$ a shortest sequence of reversals from $\pi_t$ to $\pi_0$. Then
+\begin{align*}
+    d(\pi_i) &= d(\pi_{i-1}) + 1 \\
+             \overset{(\ref{thm:1})}&{\geq} d(\pi_{i-1}) + b(\pi_i) - b(\pi_{i-1}) + c(\pi_{i-1}) - c(\pi_i) \\[0.25cm]
+    \iff & \begin{aligned}[t]
+    d(\pi_i) - (b(\pi_i) - c(\pi_i)) &\geq d(\pi_{i-1}) - (b(\pi_{i-1}) - c(\pi_{i-1})) \\
+                                     &\geq d(\pi_0) - (b(\pi_0) - c(\pi_0)) = 0
+    \end{aligned}
+\end{align*}
+Setting $i=t$, proves the theorem.
+\end{proof}
+
+Using this lower bound, \citeauthor*{Bafna1996} already describe a $\frac{7}{4}$-approximation that is not considered in more detail here.
+\citeauthor*{Christie1998} was able to formulate a lower bound in terms of the number of $2$-cycles in a maximum alternating cycle decomposition \cite{Christie1998}.
+
+\begin{lemma}
+\label{lem:2}
+Let $\mathcal{C}$ be a maximum alternating cycle decomposition of $G(\pi)$.
+Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in $\mathcal{C}$ and let $c_{3*}(\pi) = c(\pi) - c_2(\pi)$ be the remaining alternating cycles in $\mathcal{C}$.
+Then
+\begin{align*}
+    c_{3*}(\pi) \leq \frac{1}{3}(b(\pi) - 2 c_2(\pi)).
+\end{align*}
+\end{lemma}
+
+\begin{proof}
+$b(\pi)$ is the number of red edges in $G(\pi)$ and $c_2(\pi)$ is the number of red edges in $2$-cycles of $G(\pi)$.
+Therefore, $b(\pi) - 2 c_2(\pi)$ is the number of \textit{free} red edges in $G(\pi)$ not occurring in $2$-cycles.
+$3 c_{3*}(\pi)$ is a lower bound to the number of red edges in $G(\pi)$ not occurring in $2$-cycles, proving the lemma.
+\end{proof}
+
+\begin{theorem}
+\label{thm:3}
+Let $\pi \in S_n$. Then
+\begin{align*}
+    d(\pi) \geq \frac{2}{3} b(\pi) - \frac{1}{3} c_2(\pi).
+\end{align*}
+\end{theorem}
+
+\begin{proof}
+Follows immediately from theorem \ref{thm:2} and lemma \ref{lem:2}.
+\end{proof}
+
+\section{3/2-approximation}
+\label{sec:32_approximation}
+
+In this section we discuss the $3/2$-approximation proposed by \citeauthor*{Christie1998} in detail \cite{Christie1998}.
+First, we describe how a sorting sequence of reversals can be obtained from an alternating cycle decomposition of $G(\pi)$ (section \ref{sec:reversal_graph}).
+In section \ref{sec:cycle_decomposition} we then describe how to find such a cycle decomposition.
+This then allows us to prove the approximation bound (section \ref{sec:approximation_bound}).
+Lastly, we describe the full algorithm and analyze its complexity respectively (section \ref{sec:algorithm}).
+
+\subsection{Reversal graph}
+\label{sec:reversal_graph}
+
+To obtain a sorting sequence of reversals for a permutation $\pi$ given an alternating cycle decomposition $\mathcal{C}$ of its breakpoint graph $G(\pi)$, \citeauthor*{Christie1998} defines another graph.
+
+\begin{definition}
+Given an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$, the \textit{reversal graph} $R(\mathcal{C})$ is constructed as follows.
+
+For each adjacency in $\pi$ we introduce an isolated blue vertex.
+For each $m$-cycle in $\mathcal{C}$ we introduce $m$ vertices, each representing a constituting blue edge.
+Each vertex $u$ represents the reversal $\rho(u)$ that acts on the two red edges incident to the blue edge.
+In the case of a $2$-cycle, two vertices in the reversal graph represent the same reversal.
+A vertex is colored red if the represented reversal is a $1$- or a $2$-reversal.
+Otherwise a vertex is colored blue.
+
+Let $u$ be a vertex in $R(\mathcal{C})$.
+Define $l_b(u)$ and $r_b(u)$ to be the positions in $\pi$ of the leftmost and rightmost elements, respectively, that are incident to the blue edge in $G(\pi)$ represented by $u$.
+Similarly, $l_r(u)$ and $r_r(u)$ are defined to be the positions in $\pi$ of the leftmost and rightmost red edges that are adjacent to the blue edge represented by $u$.
+Hereby, the position of a red edge is the position of its leftmost incident vertex.
+Two vertices $u$ and $v$ are connected with an edge if $l_b(u) < l_b(v) < r_b(u) < r_b(v)$ or $l_b(v) < l_b(u) < r_b(v) < r_b(u)$ or $l_r(u) < l_r(v) < r_r(u) < r_r(v)$ or $l_r(v) < l_r(u) < r_r(v) < r_r(u)$.
+
+Two cycles $C$ and $D$ \textit{interleave} if the reversal graph contains vertices $u$ and $v$ arising from $C$ and $D$ respectively, such that $\{u,v\}$ is an edge in $R(\mathcal{C})$.
+Blue edges \textit{interleave} in the breakpoint graph if the vertices representing them are adjacent in $R(\mathcal{C})$.
+\end{definition}
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}
+    	\begin{pgfonlayer}{nodelayer}
+    		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+    		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
+    		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
+    		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
+    		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
+    	\end{pgfonlayer}
+    	\begin{pgfonlayer}{edgelayer}
+    		\draw (0) to (2);
+    		\draw (1) to (2);
+    	\end{pgfonlayer}
+    \end{tikzpicture}
+    \caption{Reversal graph}
+    \label{fig:reversal_graph}
+\end{figure}
+
+\begin{example}
+Consider the alternating cycle decomposition
+\begin{align*}
+    \mathcal{C} = \{(&\{1,3\},\{2,3\},\{2,4\}, \\
+                     &\{4,5\},\{2,5\},\{1,2\})\}
+\end{align*}
+of the breakpoint graph of $(0\ 1\ 3\ 4\ 2\ 5)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.
+\end{example}
+
+We observe that the only alternating cycle decomposition of $G(id)$ is $\mathcal{C} = \emptyset$.
+And therefore, the only reversal graph of the identity permutation $R(\emptyset)$ consists of $n+1$ isolated blue vertices.
+In the remainder of this chapter it our goal is to describe a way of transforming the reversal graph of a given alternating cycle decomposition into the reversal graph only consisting of isolated blue vertices with as few reversals as possible.
+
+Let $u$ be a vertex of $R(\mathcal{C})$ representing reversal $\rho(u)$.
+Denote by $\mathcal{C}_u$ the alternating cycle decomposition of $G(\pi \circ \rho(u))$ that is obtained from $\mathcal{C}$ by removing one 2-cycle or shortening one of its cycles by one red edge and one blue edge.
+
+The following lemma is given without proof.
+
+\begin{lemma}
+\label{lem:3}
+$R(\mathcal{C}_u)$ can be derived from $R(\mathcal{C})$ by making the following changes to $R(\mathcal{C})$:
+\begin{enumerate}
+    \item flip the color of every vertex adjacent to $u$;
+    \item flip the adjacency of every pair of vertices adjacent to $u$; and
+    \item if $u$ is a red vertex, turn it into an isolated blue vertex.
+\end{enumerate}
+\end{lemma}
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
+    	\begin{pgfonlayer}{nodelayer}
+    		\node [style=node] (0) at (-3, 0) {0};
+    		\node [style=node] (1) at (-1.5, 0) {1};
+    		\node [style=node] (2) at (0, 0) {2};
+    		\node [style=node] (3) at (1.5, 0) {4};
+    		\node [style=node] (4) at (3, 0) {3};
+    		\node [style=node] (5) at (4.5, 0) {5};
+    	\end{pgfonlayer}
+    	\begin{pgfonlayer}{edgelayer}
+    		\draw [style=red edge] (2) to (3);
+    		\draw [style=red edge] (4) to (5);
+    		\draw [style=blue edge, bend left] (2) to (4);
+    		\draw [style=blue edge, bend left] (3) to (5);
+    	\end{pgfonlayer}
+    \end{tikzpicture}
+    \caption{Resulting breakpoint graph}
+    \label{fig:breakpoint_graph_change}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
+    	\begin{pgfonlayer}{nodelayer}
+    		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+    		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
+    		\node [style=red node] (2) at (-1, -1) {$\{4,5\}$};
+    		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
+    		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
+    	\end{pgfonlayer}
+    	\begin{pgfonlayer}{edgelayer}
+    		\draw (0) to (2);
+    	\end{pgfonlayer}
+    \end{tikzpicture}
+    \caption{Resulting reversal graph}
+    \label{fig:reversal_graph_change}
+\end{figure}
+
+\begin{example}
+Applying the reversal represented by vertex $\{1,2\}$ in figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
+\end{example}
+
+\begin{corollary}
+A reversal $\rho(u)$ affects only vertices that are in the same connected component as $u$.
+\end{corollary}
+
+We now prove a couple of lemmas that we use later.
+
+\begin{lemma}
+\label{lem:4}
+$R(\mathcal{C})$ contains no isolated blue vertices arising from cycles in $\mathcal{C}$.
+\end{lemma}
+
+\begin{proof}
+Omitted \cite{Christie1998}.
+\end{proof}
+
+The following is a simple consequence of the above lemma.
+
+\begin{lemma}
+\label{lem:5}
+Vertices arising from an unoriented $2$-cycle of $\mathcal{C}$ must be in a connected component of $R(\mathcal{C})$ with vertices arising from at least one more alternating cycle of $\mathcal{C}$.
+\end{lemma}
+
+\begin{proof}
+Let $u$ and $v$ represent the blue edges of an unoriented $2$-cycle of $\mathcal{C}$.
+Then its two constituting blue edges do not interleave.
+Hence, $R(\mathcal{C})$ does not connect $u$ and $v$ with an edge.
+If its constituting blue edges do not interleave with blue edges of any other alternating cycle of $\mathcal{C}$, then $u$ and $v$ would be isolated blue vertices.
+$\lightning$
+\end{proof}
+
+\begin{lemma}
+\label{lem:6}
+All vertices arising from the same alternating cycle in $\mathcal{C}$ are in the same connected component of $R(\mathcal{C})$.
+\end{lemma}
+
+\begin{proof}
+Omitted \cite{Christie1998}.
+\end{proof}
+
+We call a connected component of $R(\mathcal{C})$ \textit{oriented} if it contains a red vertex or if it consists solely of an isolated blue vertex.
+
+Let $A$ be a connected component of $R(\mathcal{C})$.
+We denote by $A_u$ the subgraph of $R(\mathcal{C}_u)$ that contains all the vertices of $A$.
+A sequence of reversals that turns the connected component $A$ of $R(\mathcal{C})$ into an isolated blue vertex, is called an \textit{elemination sequence} for $A$.
+
+\begin{lemma}[\citeauthor*{Christie1998}]
+\label{lem:7}
+If a connected component $A$ of $R(\mathcal{C})$ is oriented and not an isolated blue vertex, it contains a red vertex $u$ such that $A_u$ is still oriented.
+\end{lemma}
+
+\begin{proof}
+Omitted \cite{Christie1998}.
+\end{proof}
+
+With the above lemma we are now able to find an elimination sequence for the connected components of the reversal graph.
+
+\begin{lemma}
+\label{lem:8}
+Let $A$ be an oriented connected component of $R(\mathcal{C})$ arising from $k$ different alternating cycles of $G(\pi)$.
+Then there is an elimination sequence for $A$ that contains $k$ $2$-reversals with all the other reversals being $1$-reversals.
+\end{lemma}
+
+\begin{proof}
+As $A$ is oriented, we can use lemma \ref{lem:7} to repeatedly find a red vertex $u$ in $A$ such that $A_u$ is still oriented and apply $\rho(u)$.
+By lemma \ref{lem:6}, we know that every vertex in $R(\mathcal{C})$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component.
+Therefore, every alternating cycle represented by vertices constituting $A$ eventually reduces to a $2$-cycle.
+Moreover, by lemma \ref{lem:1}, $\rho(u)$ is a $2$-reversal iff $u$ arises from an oriented $2$-cycle.
+
+\end{proof}
+
+\begin{lemma}
+\label{lem:9}
+Let $A$ be an unoriented connected component of $R(\mathcal{C})$ arising from $k$ different alternating cycles of $G(\pi)$.
+Then there is an elimination sequence for $A$ that contains one $0$-reversal, $k$ $2$-reversals, with all the other reversals being $1$-reversals.
+\end{lemma}
+
+\begin{proof}
+Let $u$ be any blue vertex of $A$, then $\rho(u)$ is a $0$-reversal. By lemma \ref{lem:3}, $A_u$ is oriented.
+Now, apply lemma \ref{lem:8} to obtain an elimination sequence for $A_u$.
+\end{proof}
+
+Using the elimination sequences we found for the connected components of $R(\mathcal{C})$ we are finally able to give an upper bound to the number of reversals we need to sort $\pi$.
+
+\begin{theorem}
+\label{thm:4}
+Let $\pi \in S_n$ and $\mathcal{C}$ be an alternating cycle decomposition of $G(\pi)$. Then
+\begin{align*}
+    d(\pi) \leq b(\pi) - |\mathcal{C}| + u(\mathcal{C})
+\end{align*}
+where $u(\mathcal{C})$ is the number of unoriented components in $R(\mathcal{C})$.
+\end{theorem}
+
+\subsection{Cycle decomposition}
+\label{sec:cycle_decomposition}
+
+In this section our goal is to find an alternating cycle decomposition of $G(\pi)$.
+Motivated by the lower bound to the reversal distance in terms of $2$-cycles, we try to maximize the number of $2$-cycles in our cycle decomposition \cite{Christie1998}.
+
+To find a large number of $2$-cycles in $G(\pi)$, \citeauthor*{Christie1998} constructs a new graph.
+
+\begin{definition}
+Given the permutation $\pi$, we construct the \textit{matching graph} $F(\pi)$ with vertices representing red edges in $G(\pi)$ and vertices $u, v$ being adjacent if they form a $2$-cycle in $G(\pi)$.
+\end{definition}
+
+We now observe that $2$-cycles of $G(\pi)$ are represented by edges in $F(\pi)$.
+We then find a maximum cardinality matching $M$ of $F(\pi)$.
+We call a $2$-cycle of $G(\pi)$ \textit{selected} if its corresponding edge of $F(\pi)$ is part of the matching $M$.
+Note, that we haven't yet found the set of edge-disjoint $2$-cycles that we need to construct an alternating cycle decomposition.
+This is because the matching ensures that all selected $2$-cycles are edge-disjoint on their red edges, but does not ensure that they are also edge-disjoint on their blue edges.
+To find a large set of edge-disjoint $2$-cycles, \citeauthor*{Christie1998} introduces another graph.
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}
+    	\begin{pgfonlayer}{nodelayer}
+    		\node [style=node] (0) at (-4.5, 0) {};
+    		\node [style=node] (1) at (-3, 0) {};
+    		\node [style=node] (2) at (-1.5, 0) {};
+    		\node [style=node] (3) at (0, 0) {};
+    		\node [style=node] (4) at (1.5, 0) {};
+    		\node [style=node] (5) at (3, 0) {};
+    		\node [style=node] (6) at (4.5, 0) {};
+    		\node [style=node] (7) at (6, 0) {};
+    	\end{pgfonlayer}
+    	\begin{pgfonlayer}{edgelayer}
+    		\draw [style=red edge] (0) to (1);
+    		\draw [style=red edge] (1) to (2);
+    		\draw [style=red edge] (2) to (3);
+    		\draw [style=red edge] (4) to (5);
+    		\draw [style=red edge] (5) to (6);
+    		\draw [style=red edge] (6) to (7);
+    		\draw [style=blue edge, bend left] (3) to (4);
+    		\draw [style=blue edge, bend left] (2) to (5);
+    		\draw [style=blue edge, bend left] (1) to (6);
+    		\draw [style=blue edge, bend left] (0) to (7);
+    	\end{pgfonlayer}
+    \end{tikzpicture}
+    \caption{Ladder}
+    \label{fig:ladder}
+\end{figure}
+
+\begin{definition}
+Given a matching $M$, we construct the \textit{ladder graph} $L(M)$.
+$L(M)$ has a vertex for every $2$-cycle selected in $M$.
+Vertices form connected components when the represented $2$-cycles share a blue edge in $G(\pi)$.
+Such connected components of $L(M)$ are called \textit{ladders}.
+Figure \ref{fig:ladder} shows part of a breakpoint graph with $2$-cycles that form a ladder.
+
+A selected $2$-cycle is called \textit{independent} if it is not part of a ladder. Otherwise it is called a \textit{ladder $2$-cycle}.
+\end{definition}
+
+As each $2$-cycle can share at most two blue edges with other $2$-cycles, the connected components of $L(M)$ are simple paths.
+Therefore, the set of $2$-cycles consisting of every other ladder $2$-cycle is edge-disjoint on all edges, resulting in the following theorem.
+
+\begin{theorem}
+\label{thm:5}
+Given a maximum cardinality matching $M$ of $F(\pi)$ it is possible to find an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$ that contains at least $\left\lceil \frac{y}{2} \right\rceil$ ladder $2$-cycles and $z$ independent $2$-cycles.
+\end{theorem}
+
+\begin{proof}
+Let $\mathcal{C}$ contain all independent $2$-cycles from $L(M)$ and every other alternate cycle of each ladder.
+Obtain the rest of $\mathcal{C}$ by adding any alternating cycle decomposition of the remaining unused edges of $G(\pi)$.
+\end{proof}
+
+Another lemma, \citeauthor*{Christie1998} shows, is given without proof.
+
+\begin{lemma}
+\label{lem:10}
+Let $C$ be an unoriented ladder $2$-cycle in $\mathcal{C}$.
+Then the vertices of $R(\mathcal{C})$ that represent $C$ are part of a component that contains vertices representing an alternating cycle that is not a selected $2$-cycle.
+\end{lemma}
+
+\subsection{Approximation bound}
+\label{sec:approximation_bound}
+
+By theorem \ref{thm:3}, an algorithm that finds a sorting sequence of reversals of at most length $b(\pi) - \frac{1}{2} c_2(\pi)$ achieves an approximation bound of $\frac{3}{2}$.
+
+\begin{theorem}
+\label{thm:6}
+Let $\pi \in S_n$. Then
+\begin{align*}
+    d(\pi) \leq b(\pi) - \frac{1}{2} c_2(\pi).
+\end{align*}
+\end{theorem}
+
+\begin{proof}
+Using theorem \ref{thm:5}, first find an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$ with at least $\left\lceil \frac{y}{2} \right\rceil$ ladder $2$-cycles and $z$ independent $2$-cycles.
+
+Let $k$ be the number of $2$-cycles in oriented connected components of $R(\mathcal{C})$.
+Let $u$ be the number of unoriented connected components in $R(\mathcal{C})$ that include $l$ selected $2$-cycles and that, by lemma \ref{lem:10}, contain vertices representing remaining unselected $2$-cycles.
+Let $v$ be the number of remaining unoriented connected components consisting only of vertices representing $m$ independent selected $2$-cycles.
+
+Note that $\mathcal{C}$ has at least $k+l+u+m$ alternating cycles and $R(\mathcal{C})$ has $u+v$ unoriented connected components.
+
+By theorem \ref{thm:4}, we can now sort $\pi$ using at least $k+l+u+m$ $2$-reversals and only $u+v$ $0$-reversals. Therefore
+\begin{align*}
+    d(\pi) &\leq b(\pi) - k - l - u - m + u + v \\
+           &= b(\pi) - k - l - m + v.
+\end{align*}
+All that is left to show is $k+l+m-v \geq \frac{1}{2} c_2(\pi)$. We know:
+\begin{enumerate}
+    \item $l+m$ is the number of selected $2$-cycles in unoriented connected components of $R(\mathcal{C})$.
+$k$ is the total number of $2$-cycles in oriented connected components of $R(\mathcal{C})$.
+As $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles in $\mathcal{C}$, $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ holds.
+    \item As each connected components of $R(\mathcal{C})$ counted by $v$ represents at least one unoriented $2$-cycle, lemma \ref{lem:5} implies that such a connected component must also represent at least one more alternating cycle in $\mathcal{C}$.
+As the connected component only represents independent $2$-cycles, we know that it must represent at least two independent $2$-cycles.
+As $z$ is the number of independent $2$-cycles, $v \leq \left\lfloor \frac{z}{2} \right\rfloor$ holds.
+    \item As the set of selected $2$-cycles is not necessarily edge-disjoint, $|M| = y+z \geq c_2(\pi)$.
+\end{enumerate}
+And therefore
+\begin{align*}
+    k+l+m-v &\geq \left\lceil \frac{y}{2} \right\rceil + z - v                                          & (1) \\
+            &\geq \left\lceil \frac{y}{2} \right\rceil + z - \left\lfloor \frac{z}{2} \right\rfloor     & (2) \\
+            &= \left\lceil \frac{y}{2} \right\rceil + \left\lceil \frac{z}{2} \right\rceil \\
+            &\geq \frac{1}{2} c_2(\pi).                                                                 & (3)
+\end{align*}
+\end{proof}
+
+\subsection{Algorithm}
+\label{sec:algorithm}
+
+The approximation algorithm with a performance ratio of $\frac{3}{2}$, is outlined in the following.
+
+\begin{algorithm}
+    \SetKwInOut{Input}{Input}
+    \Input{permutation $\pi \in S_n$}
+    construct the matching graph $F(\pi)$\;
+    find a maximum cardinality matching $M$ for this graph\;
+    construct alternating cycle decomposition $\mathcal{C}$ using this matching\;
+    construct the reversal graph $R(\mathcal{C})$\;
+    find an elimination sequence for $R(\mathcal{C})$\;
+    \caption{Approximation algorithm \cite{Christie1998}}
+\end{algorithm}
+
+The matching graph can be constructed in $O(n)$ time.
+The maximum cardinality matching can be obtained from $F(\pi)$ using the algorithm of Micali and Vazirani in $O(|V||E|^{\frac{1}{2}})$ time \cite{Micali1980} where $|V|$ and $|E|$ are the number of vertices and edges respectively.
+Therefore, it is possible to find such a matching in $O(n^{\frac{3}{2}})$ time.
+The alternating cycle decomposition can be obtained from the matching in $O(n)$ time.
+The reversal graph can be constructed in $O(n^2)$ time.
+The algorithm to find an elimination sequence of $R(\mathcal{C})$ is outlined below and terminates in $O(n^4)$ time.
+So the overall time-complexity of the algorithm is $O(n^4)$ \cite{Christie1998}.
+Using results from \citeauthor*{Kaplan1997}, the time-complexity can be improved to $O(n^2)$ \cite{Kaplan1997}.
+
+\begin{algorithm}
+    \SetKwInOut{Input}{Input}
+    \Input{reversal graph $R(\mathcal{C})$}
+    for each unoriented connected component, apply a reversal represented by a blue vertex\;
+    \While{not sorted $\pi$}{
+        u := first red vertex\;
+        found := false\;
+        \While{not found}{
+            apply $\rho(u)$ to get $\mathcal{C}_u$\;
+            generate $R(\mathcal{C}_u)$\;
+            count unoriented connected components in $R(\mathcal{C}_u)$\;
+            \eIf{there are no unoriented connected components}{
+                found := true\;
+            }{
+                u := next red vertex\;
+            }
+        }
+    }
+    \caption{Finding an elimination sequence \cite{Christie1998}}
+\end{algorithm}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+We have seen multiple layers of abstraction that allowed us to describe a $3/2$-approximation.
+As was mentioned in the introduction, a better but more complex approximation was obtained later by \citeauthor*{Berman2001}.
+
+An interesting property of the discussed $3/2$-approximation is that when constructing the alternating cycle decomposition, any cycle decomposition of the edges that are not part of $2$-cycles will suffice \cite{Christie1998}.
+
 \printbibliography

### 2020-07-11 paper.tex
update paper

@@
 \def\code#1{\texttt{\frenchspacing#1}}
-\def\padding{\vspace{3em}}
+\def\padding{\vspace{2em}}
 
@@
     A reversal is a transformation, reversing the order of elements of a permutation.
-Sorting by Reversals is the problem of finding the shortest sequence of reversals transforming a permutation to the identity.
+Sorting by reversals is the problem of finding the shortest sequence of reversals transforming a permutation to the identity.
 The length of such a sequence is said to be the reversal distance of that permutation.
 In this paper we introduce fundamental concepts used in the literature on this topic.
-In particular, the fundamental relationship between the reversal distance and the number of alternating cycles in a cycle decomposition of a suitably-defined bi-colored graph.
-Then a $3/2$-approximation is discussed in detail.
+In particular, we introduce the fundamental relationship between the reversal distance and the number of alternating cycles in a cycle decomposition of a suitably-defined bi-colored graph.
+Then, a $3/2$-approximation is discussed in detail.
 We also reference further related research.
@@
 
-In computational biology a fundamental problem is to identify the evolutionary similarity between genome sequences.
+In computational biology a fundamental problem is to determine the evolutionary similarity between genome sequences.
 By far the most common mutation on genomes is the inversion.
 A genome can be represented by a permutation.
-An inversion then simply is the reversal of the substring of a permutation \cite{Kececioglu1995}.
+An inversion then simply is the reversal of a substring of a permutation \cite{Kececioglu1995}.
 
@@
 The identity permutation $id = (1\ \dots\ n)$ is the permutation mapping every input to itself.
-A \textit{reversal} $\rho(i,j)$ inverses the substring from $\pi_i$ to $\pi_j$ of a permutation.
+A \textit{reversal} $\rho(i,j) \in S_n$ inverses the substring from $\pi_i$ to $\pi_j$ of a permutation.
 It is represented by the permutation
@@
 \begin{definition}
-Given two permutations $\sigma$ and $\tau$, the \textit{reversal distance problem} is the problem of finding the shortest sequence of reversals transforming $\sigma$ into $\tau$.
-More formally, we want to find reversals $\rho_1, \dots, \rho_d$ such that
+Given two permutations $\sigma, \tau \in S_n$, the \textit{reversal distance problem} is the problem of finding the shortest sequence of reversals transforming $\sigma$ into $\tau$.
+More formally, we want to find reversals $\rho_1, \dots, \rho_d \in S_n$ such that
 \begin{align*}
@@
 \begin{definition}
-\textit{Sorting by Reversals (MIN-SBR)} then is the problem of determining the reversal distance between some permutation $\pi$ and $id$.
+\textit{Sorting by reversals (MIN-SBR)} then is the problem of determining the reversal distance between some permutation $\pi \in S_n$ and $id$.
 The reversal distance is denoted by $d(\pi)$.
@@
 In this paper we give an overview of approximation algorithms for MIN-SBR.
-In section \ref{sec:breakpoints} we first introduce the notion of breakpoints introduced by \citeauthor*{Kececioglu1995} that allow us to formulate a first lower bound for the reversal distance of $\pi$ and lead to a $2$-approximation \cite{Kececioglu1995}.
+We first introduce the notion of breakpoints given by \citeauthor*{Kececioglu1995} allowing us to formulate a first lower bound for the reversal distance of $\pi$ and leading to a $2$-approximation \cite{Kececioglu1995} (section \ref{sec:breakpoints}).
 In section \ref{sec:breakpoint_graph} we define a suitable bi-colored graph first introduced by \citeauthor*{Bafna1996} to find a tighter lower bound \cite{Bafna1996}.
-In this section, we make a fundamental observation relating the reversal distance to the number of alternating cycles in a cycle decomposition of such a graph.
+In doing so, we make a fundamental observation relating the reversal distance to the number of alternating cycles in a cycle decomposition of such a graph.
 Lastly, we consider a $3/2$-approximation described by \citeauthor{Christie1998} \cite{Christie1998} (Section \ref{sec:32_approximation}).
@@
 
-Now we consider the relationship between two consecutive elements of a permutation.
+A very fundamental way of approaching the problem of sorting by reversals is to consider the relationship between two consecutive elements of a permutation.
 
@@
 
-In an effort to make the identity permutation the only permutation without any breakpoints, we extend our definition of permutations (and the symmetric group) by expanding permutations with one fixed initial $0$ and one fixed trailing $n+1$.
+In an effort to make the identity permutation the only permutation without any breakpoints, we extend our definition of permutations (and the symmetric group) by expanding permutations with one initial $0$ and one trailing $n+1$.
 In the following we only consider such expanded permutations $\pi = (0\ \pi_1\ \dots\ \pi_n\ n+1) \in S_n$ where $\pi_0 = 0$ and $\pi_{n+1} = n+1$.
@@
 \begin{corollary}
-Using the previous two observations, \citeauthor*{Kececioglu1995} found a first fundamental lower bound to the reversal distance of a permutation $\pi$ \cite{Kececioglu1995}.
+Using the previous two observations, \citeauthor*{Kececioglu1995} found a first fundamental lower bound to the reversal distance of a permutation $\pi \in S_n$ \cite{Kececioglu1995}.
 \begin{align*}
@@
 
-Based on this first lower bound, \citeauthor*{Kececioglu1995} proposed a simple $2$-approximation running in $O(n^2)$ \cite{Kececioglu1995}.
+Based on this first lower bound, \citeauthor*{Kececioglu1995} proposed a simple $2$-approximation running in $O(n^2)$ time \cite{Kececioglu1995}.
 
@@
 
-We now try to find a tighter lower bound by introducing the breakpoint graph of a permutation.
+We now try to find a tighter lower bound by introducing the breakpoint graph of a permutation first defined by \citeauthor*{Bafna1996}.
 
@@
 
-The red edges in the breakpoint graph identify the elements of $\pi$ that have to be split apart and the blue edges identify the elements of $\pi$ that have to be brought together (the missing adjacencies) in the process of sorting the permutation $\pi$.
+The red edges in the breakpoint graph identify the elements of $\pi$ that have to be split apart and the blue edges identify the elements of $\pi$ that have to be brought together in the process of sorting the permutation $\pi$ (the missing adjacencies).
 Figure \ref{fig:breakpoint_graph} shows the breakpoint graph of the permutation $(0\ 1\ 3\ 4\ 2\ 5)$.
 
-An \textit{alternating cycle decomposition} of $G(\pi)$ is a collection of edge-disjoint alternating cycles, such that every edge of $G(\pi)$ is contained in exactly one cycle of the collection.
+An \textit{alternating cycle decomposition} of $G(\pi)$ is a set of edge-disjoint alternating cycles, such that every edge of $G(\pi)$ is contained in exactly one cycle of the set.
 We observe that for every breakpoint, a vertice in $G(\pi)$ must have a missing adjacency.
-And conversely, for every missing adjacency a vertex in $G(\pi)$ must have a breakpoint.
+And conversely, for every missing adjacency a vertex in $G(\pi)$ must form a breakpoint.
 Therefore, each vertex in our breakpoint graph has an equal number of incident red and blue edges.
@@
 
-The main result we prove in this section, relates the maximum number of alternating cycles of any cycle decomposition of $G(\pi)$ to the reversal distance of $\pi$.
+The main result we prove in this section, relates the maximum number of alternating cycles of any cycle decomposition of $G(\pi)$ to the reversal distance of $\pi$.\padding
 
@@
 Identify with $k$ and $l$ the vertices incident to the first red edge and with $m$ and $n$ the vertices incident to the second red edge.
-Now, because $\rho$ is removes both red edges from the breakpoint graph, we know that either $k \sim m$ and $l \sim n$ or $k \sim n$ and $l \sim m$.
+Now, because $\rho$ removes both red edges from the breakpoint graph, we know that either $k \sim m$ and $l \sim n$ or $k \sim n$ and $l \sim m$.
 Thus, forming a $2$-cycle in $G(\pi)$.
@@
 Using this lower bound, \citeauthor*{Bafna1996} already describe a $\frac{7}{4}$-approximation that is not considered in more detail here.
-\citeauthor*{Christie1998} was able to formulate a lower bound in terms of the number of $2$-cycles in a maximum alternating cycle decomposition \cite{Christie1998}.
+\citeauthor*{Christie1998} formulated a lower bound in terms of the number of $2$-cycles in a maximum alternating cycle decomposition \cite{Christie1998}.
 
@@
 This then allows us to prove the approximation bound (section \ref{sec:approximation_bound}).
-Lastly, we describe the full algorithm and analyze its complexity respectively (section \ref{sec:algorithm}).
+Lastly, we describe the full algorithm and analyze its complexity (section \ref{sec:algorithm}).
 
@@
 
-To obtain a sorting sequence of reversals for a permutation $\pi$ given an alternating cycle decomposition $\mathcal{C}$ of its breakpoint graph $G(\pi)$, \citeauthor*{Christie1998} defines another graph.
+To obtain a sorting sequence of reversals for a permutation $\pi$ given an alternating cycle decomposition $\mathcal{C}$ of its breakpoint graph $G(\pi)$, \citeauthor*{Christie1998} defines a new graph.
 
@@
 Similarly, $l_r(u)$ and $r_r(u)$ are defined to be the positions in $\pi$ of the leftmost and rightmost red edges that are adjacent to the blue edge represented by $u$.
-Hereby, the position of a red edge is the position of its leftmost incident vertex.
+The position of a red edge is the position of its leftmost incident vertex.
 Two vertices $u$ and $v$ are connected with an edge if $l_b(u) < l_b(v) < r_b(u) < r_b(v)$ or $l_b(v) < l_b(u) < r_b(v) < r_b(u)$ or $l_r(u) < l_r(v) < r_r(u) < r_r(v)$ or $l_r(v) < l_r(u) < r_r(v) < r_r(u)$.
@@
 And therefore, the only reversal graph of the identity permutation $R(\emptyset)$ consists of $n+1$ isolated blue vertices.
-In the remainder of this chapter it our goal is to describe a way of transforming the reversal graph of a given alternating cycle decomposition into the reversal graph only consisting of isolated blue vertices with as few reversals as possible.
+In the remainder of this section, our goal is to describe a way of transforming the reversal graph of a given alternating cycle decomposition into the reversal graph only consisting of isolated blue vertices with as few reversals as possible.
+For that purpose, we first examine the effects of single reversals on the reversal graph.\padding
 
@@
 
-We now prove a couple of lemmas that we use later.
+We now consider a couple of lemmas that will help use prove the approximation bound later.
 
@@
 \label{lem:4}
-$R(\mathcal{C})$ contains no isolated blue vertices arising from cycles in $\mathcal{C}$.
+$R(\mathcal{C})$ contains no isolated blue vertices arising from alternating cycles in $\mathcal{C}$.
 \end{lemma}
@@
 
-The following is a simple consequence of the above lemma.
+The following lemma is a simple consequence of the above lemma.
 
@@
 Omitted \cite{Christie1998}.
-\end{proof}
+\end{proof}\padding
 
@@
 We denote by $A_u$ the subgraph of $R(\mathcal{C}_u)$ that contains all the vertices of $A$.
-A sequence of reversals that turns the connected component $A$ of $R(\mathcal{C})$ into an isolated blue vertex, is called an \textit{elemination sequence} for $A$.
+A sequence of reversals that turns the connected component $A$ of $R(\mathcal{C})$ into an isolated blue vertex, is called an \textit{elemination sequence} of $A$.
 
-\begin{lemma}[\citeauthor*{Christie1998}]
+\begin{lemma}
 \label{lem:7}
@@
 By lemma \ref{lem:6}, we know that every vertex in $R(\mathcal{C})$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component.
-Therefore, every alternating cycle represented by vertices constituting $A$ eventually reduces to a $2$-cycle.
+Therefore, every alternating cycle represented by vertices in $A$ eventually reduces to a $2$-cycle.
 Moreover, by lemma \ref{lem:1}, $\rho(u)$ is a $2$-reversal iff $u$ arises from an oriented $2$-cycle.
-
 \end{proof}
@@
 Such connected components of $L(M)$ are called \textit{ladders}.
-Figure \ref{fig:ladder} shows part of a breakpoint graph with $2$-cycles that form a ladder.
+Figure \ref{fig:ladder} shows a part of a breakpoint graph with $2$-cycles that form a ladder.
 
@@
 \label{thm:6}
-Let $\pi \in S_n$. Then
-\begin{align*}
-    d(\pi) \leq b(\pi) - \frac{1}{2} c_2(\pi).
-\end{align*}
+Let $\pi \in S_n$. We are able to find a sorting sequence of at most $b(\pi) - \frac{1}{2} c_2(\pi)$ reversals.
 \end{theorem}
@@
 As $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles in $\mathcal{C}$, $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ holds.
-    \item As each connected components of $R(\mathcal{C})$ counted by $v$ represents at least one unoriented $2$-cycle, lemma \ref{lem:5} implies that such a connected component must also represent at least one more alternating cycle in $\mathcal{C}$.
+    \item As each connected components of $R(\mathcal{C})$ counted by $v$ represents at least one unoriented $2$-cycle, lemma \ref{lem:5} implies that such a connected component must represent at least two alternating cycles in $\mathcal{C}$.
 As the connected component only represents independent $2$-cycles, we know that it must represent at least two independent $2$-cycles.
 As $z$ is the number of independent $2$-cycles, $v \leq \left\lfloor \frac{z}{2} \right\rfloor$ holds.
-    \item As the set of selected $2$-cycles is not necessarily edge-disjoint, $|M| = y+z \geq c_2(\pi)$.
+    \item We also have that the selected $2$-cycles in $M$ are not necessarily edge-disjoint on their blue edges, yielding $|M| = y+z \geq c_2(\pi)$.
 \end{enumerate}
@@
 
-The approximation algorithm with a performance ratio of $\frac{3}{2}$, is outlined in the following.
+The approximation algorithm with a performance ratio of $\frac{3}{2}$, is outlined in figure \ref{alg:approximation}.
 
-\begin{algorithm}
+\begin{figure}[htb]
+    \begin{algorithm}[H]
         \SetKwInOut{Input}{Input}
@@
     \end{algorithm}
+    \caption{}
+    \label{alg:approximation}
+\end{figure}
 
@@
 The reversal graph can be constructed in $O(n^2)$ time.
-The algorithm to find an elimination sequence of $R(\mathcal{C})$ is outlined below and terminates in $O(n^4)$ time.
+The algorithm to find an elimination sequence of $R(\mathcal{C})$ is outlined in figure \ref{alg:elimination_sequence} and runs in $O(n^4)$ time.
 So the overall time-complexity of the algorithm is $O(n^4)$ \cite{Christie1998}.
-Using results from \citeauthor*{Kaplan1997}, the time-complexity can be improved to $O(n^2)$ \cite{Kaplan1997}.
 
-\begin{algorithm}
+Using results from \citeauthor*{Kaplan1997}, the time-complexity of finding an elimination sequence and therefore the time-complexity of the entire algorithm can be improved to $O(n^2)$ \cite{Kaplan1997}.
+
+\begin{figure}[htb]
+    \begin{algorithm}[H]
         \SetKwInOut{Input}{Input}
@@
     \end{algorithm}
+    \caption{}
+    \label{alg:elimination_sequence}
+\end{figure}
 

### 2020-07-11 paper.tex
update example in paper

@@
     \begin{align*}
-        \pi = (1\ \mid\ 3\ 4\ \mid\ 2)
+        \pi = (7\ \mid\ 5\ 6\ \mid\ 3\ 2\ \mid\ 4\ \mid\ 1)
     \end{align*}
@@
     	\begin{pgfonlayer}{nodelayer}
-    		\node [style=node] (0) at (-3, 0) {0};
-    		\node [style=node] (1) at (-1.5, 0) {1};
-    		\node [style=node] (2) at (0, 0) {3};
-    		\node [style=node] (3) at (1.5, 0) {4};
-    		\node [style=node] (4) at (3, 0) {2};
-    		\node [style=node] (5) at (4.5, 0) {5};
+    		\node [style=node] (0) at (-6, 0) {0};
+    		\node [style=node] (1) at (-4.5, 0) {7};
+    		\node [style=node] (2) at (-3, 0) {5};
+    		\node [style=node] (3) at (-1.5, 0) {6};
+    		\node [style=node] (4) at (0, 0) {3};
+    		\node [style=node] (5) at (1.5, 0) {2};
+    		\node [style=node] (6) at (3, 0) {4};
+    		\node [style=node] (7) at (4.5, 0) {1};
+    		\node [style=node] (8) at (6, 0) {8};
     	\end{pgfonlayer}
     	\begin{pgfonlayer}{edgelayer}
+    		\draw [style=red edge] (0) to (1);
     		\draw [style=red edge] (1) to (2);
     		\draw [style=red edge] (3) to (4);
-    		\draw [style=red edge] (4) to (5);
-    		\draw [style=blue edge, bend right] (1) to (4);
-    		\draw [style=blue edge, bend left] (2) to (4);
-    		\draw [style=blue edge, bend left] (3) to (5);
+    		\draw [style=red edge] (5) to (6);
+    		\draw [style=red edge] (6) to (7);
+    		\draw [style=red edge] (7) to (8);
+    		\draw [style=blue edge, bend left] (0) to (7);
+    		\draw [style=blue edge, bend left] (1) to (3);
+    		\draw [style=blue edge, bend left] (4) to (6);
+    		\draw [style=blue edge, bend right] (1) to (8);
+    		\draw [style=blue edge, bend right] (2) to (6);
+    		\draw [style=blue edge, bend right] (5) to (7);
     	\end{pgfonlayer}
@@
 The red edges in the breakpoint graph identify the elements of $\pi$ that have to be split apart and the blue edges identify the elements of $\pi$ that have to be brought together in the process of sorting the permutation $\pi$ (the missing adjacencies).
-Figure \ref{fig:breakpoint_graph} shows the breakpoint graph of the permutation $(0\ 1\ 3\ 4\ 2\ 5)$.
+Figure \ref{fig:breakpoint_graph} shows the breakpoint graph of the permutation $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$.
 
@@
 An alternating cycle of $G(\pi)$ is called \textit{undirected} if it is not directed.
-For example, in figure \ref{fig:breakpoint_graph} the alternating cycle $(\{1,3\},\{2,3\},\{2,5\},\{4,5\},\{2,4\},\{1,2\})$ is directed with respect to $(0\ 1\ 3\ 4\ 2\ 5)$, whereas the alternating cycle $(\{1,3\},\{2,3\},\{2,4\},\{4,5\},\{2,5\},\{1,2\})$ is undirected with respect to $(0\ 1\ 3\ 4\ 2\ 5)$.
+For example, in figure \ref{fig:breakpoint_graph} the alternating cycle $(\{0,7\},\{6,7\},\{3,6\},\{3,4\},\{1,4\},\{0,1\})$ is directed with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$, whereas the alternating cycle $(\{5,7\},\{4,5\},\{2,4\},\{1,2\},\{1,8\},\{7,8\})$ is undirected with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$.
 \citeauthor*{Caprara1997} argues that an alternating cycle of $G(\pi)$ is oriented if and only if it is undirected \cite{Caprara1997}.
@@
     	\begin{pgfonlayer}{nodelayer}
-    		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+    		\node [style=red node] (0) at (-4, 1) {$\{4,5\}$};
     		\node [style=red node] (1) at (-2, 1) {$\{1,2\}$};
-    		\node [style=blue node] (2) at (-1, -1) {$\{4,5\}$};
-    		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
-    		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
+    		\node [style=blue node] (2) at (0, 1) {$\{7,8\}$};
+    		\node [style=blue node] (3) at (2, 1) {$\{5,6\}$};
+    		\node [style=blue node] (4) at (-4, -1) {$\{6,7\}$};
+    		\node [style=blue node] (5) at (-2, -1) {$\{3,4\}$};
+    		\node [style=blue node] (6) at (0, -1) {$\{0,1\}$};
+    		\node [style=blue node] (7) at (2, -1) {$\{2,3\}$};
     	\end{pgfonlayer}
     	\begin{pgfonlayer}{edgelayer}
-    		\draw (0) to (2);
-    		\draw (1) to (2);
+    		\draw (0) to (4);
+    		\draw (0) to (1);
+    		\draw (0) to (5);
+    		\draw (1) to (5);
+    		\draw (1) to (6);
+    		\draw (2) to (4);
+    		\draw (2) to (6);
     	\end{pgfonlayer}
@@
 \begin{align*}
-    \mathcal{C} = \{(&\{1,3\},\{2,3\},\{2,4\}, \\
-                     &\{4,5\},\{2,5\},\{1,2\})\}
+    \mathcal{C} = \{&(\{0,7\},\{6,7\},\{3,6\},\{3,4\},\{1,4\},\{0,1\}), \\
+                    &(\{5,7\},\{4,5\},\{2,4\},\{1,2\},\{1,8\},\{7,8\})\}
 \end{align*}
-of the breakpoint graph of $(0\ 1\ 3\ 4\ 2\ 5)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.
+of the breakpoint graph of $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.
 \end{example}
@@
     \centering
-    \begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
+    \begin{tikzpicture}
     	\begin{pgfonlayer}{nodelayer}
-    		\node [style=node] (0) at (-3, 0) {0};
-    		\node [style=node] (1) at (-1.5, 0) {1};
-    		\node [style=node] (2) at (0, 0) {2};
-    		\node [style=node] (3) at (1.5, 0) {4};
-    		\node [style=node] (4) at (3, 0) {3};
-    		\node [style=node] (5) at (4.5, 0) {5};
+    		\node [style=node] (0) at (-6, 0) {0};
+    		\node [style=node] (1) at (-4.5, 0) {7};
+    		\node [style=node] (2) at (-3, 0) {2};
+    		\node [style=node] (3) at (-1.5, 0) {3};
+    		\node [style=node] (4) at (0, 0) {6};
+    		\node [style=node] (5) at (1.5, 0) {5};
+    		\node [style=node] (6) at (3, 0) {4};
+    		\node [style=node] (7) at (4.5, 0) {1};
+    		\node [style=node] (8) at (6, 0) {8};
     	\end{pgfonlayer}
     	\begin{pgfonlayer}{edgelayer}
-    		\draw [style=red edge] (2) to (3);
-    		\draw [style=red edge] (4) to (5);
-    		\draw [style=blue edge, bend left] (2) to (4);
-    		\draw [style=blue edge, bend left] (3) to (5);
+    		\draw [style=red edge] (0) to (1);
+    		\draw [style=red edge] (1) to (2);
+    		\draw [style=red edge] (3) to (4);
+    		\draw [style=red edge] (6) to (7);
+    		\draw [style=red edge] (7) to (8);
+    		\draw [style=blue edge, bend left] (0) to (7);
+    		\draw [style=blue edge, bend left] (1) to (4);
+    		\draw [style=blue edge, bend left] (3) to (6);
+    		\draw [style=blue edge, bend right] (1) to (8);
+    		\draw [style=blue edge, bend right] (2) to (7);
     	\end{pgfonlayer}
@@
     \centering
-    \begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
+    \begin{tikzpicture}
     	\begin{pgfonlayer}{nodelayer}
-    		\node [style=red node] (0) at (0, 1) {$\{2,3\}$};
+    		\node [style=blue node] (0) at (-4, 1) {$\{4,5\}$};
     		\node [style=blue node] (1) at (-2, 1) {$\{1,2\}$};
-    		\node [style=red node] (2) at (-1, -1) {$\{4,5\}$};
-    		\node [style=blue node] (3) at (2, -1) {$\{3,4\}$};
-    		\node [style=blue node] (4) at (2, 1) {$\{0,1\}$};
+    		\node [style=blue node] (2) at (0, 1) {$\{7,8\}$};
+    		\node [style=blue node] (3) at (2, 1) {$\{5,6\}$};
+    		\node [style=red node] (4) at (-4, -1) {$\{6,7\}$};
+    		\node [style=red node] (5) at (-2, -1) {$\{3,4\}$};
+    		\node [style=blue node] (6) at (0, -1) {$\{0,1\}$};
+    		\node [style=blue node] (7) at (2, -1) {$\{2,3\}$};
     	\end{pgfonlayer}
     	\begin{pgfonlayer}{edgelayer}
-    		\draw (0) to (2);
+    		\draw (1) to (4);
+    		\draw (1) to (6);
+    		\draw (2) to (4);
+    		\draw (2) to (6);
+    		\draw (4) to (5);
     	\end{pgfonlayer}
@@
 \begin{example}
-Applying the reversal represented by vertex $\{1,2\}$ in figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
+Applying the reversal represented by vertex $\{4,5\}$ in figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
 \end{example}
@@
 A reversal $\rho(u)$ affects only vertices that are in the same connected component as $u$.
-\end{corollary}
+\end{corollary}\padding
 

### 2020-07-11 paper.tex
update paper

@@
 Let $\mathcal{C}$ be a maximum alternating cycle decomposition of $G(\pi)$.
-Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in $\mathcal{C}$ and let $c_{3*}(\pi) = c(\pi) - c_2(\pi)$ be the remaining alternating cycles in $\mathcal{C}$.
+Let $c_2(\pi)$ be the minimum number of alternating $2$-cycles in any such $\mathcal{C}$ and let $c_{3*}(\pi) = c(\pi) - c_2(\pi)$ be the remaining alternating cycles in $\mathcal{C}$.
 Then
@@
 Therefore, $b(\pi) - 2 c_2(\pi)$ is the number of \textit{free} red edges in $G(\pi)$ not occurring in $2$-cycles.
-$3 c_{3*}(\pi)$ is a lower bound to the number of red edges in $G(\pi)$ not occurring in $2$-cycles, proving the lemma.
+$3 c_{3*}(\pi)$ is a lower bound to the number of red edges in $G(\pi)$ not occurring in $2$-cycles as every such cycle has at least three constituting red edges, proving the lemma.
 \end{proof}

### 2020-07-12 paper.tex
update paper conclusion

@@
 
-We have seen multiple layers of abstraction that allowed us to describe a $3/2$-approximation.
-As was mentioned in the introduction, a better but more complex approximation was obtained later by \citeauthor*{Berman2001}.
+We have seen the fundamental relationship between the reversal distance of a permutation and its number of breakpoints as well as the maximum number of alternating cycles of a cycle decomposition of its breakpoint graph as was first discussed by \citeauthor*{Bafna1996}.
+Exploiting this relationship, \citeauthor*{Christie1998} described a $3/2$-approximation we discussed in detail.
+A better but more complex approximation was obtained later by \citeauthor*{Berman2001}.
 
-An interesting property of the discussed $3/2$-approximation is that when constructing the alternating cycle decomposition, any cycle decomposition of the edges that are not part of $2$-cycles will suffice \cite{Christie1998}.
+An interesting property of the described $3/2$-approximation is that when constructing the alternating cycle decomposition, any cycle decomposition of the edges that are not part of $2$-cycles will suffice \cite{Christie1998}.
 

### 2020-07-12 paper.tex
update paper

@@
 \newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}{Example}[theorem]
 \theoremstyle{remark}
@@
     \begin{align*}
-        \pi = (7\ \mid\ 5\ 6\ \mid\ 3\ 2\ \mid\ 4\ \mid\ 1)
+        (7\ \mid\ 5\ 6\ \mid\ 3\ 2\ \mid\ 4\ \mid\ 1)
     \end{align*}
@@
 
-Using this lower bound, \citeauthor*{Bafna1996} already describe a $\frac{7}{4}$-approximation that is not considered in more detail here.
+Using this lower bound, \citeauthor*{Bafna1996} already describe a $7/4$-approximation that is not considered in more detail here.
 \citeauthor*{Christie1998} formulated a lower bound in terms of the number of $2$-cycles in a maximum alternating cycle decomposition \cite{Christie1998}.
@@
 $b(\pi)$ is the number of red edges in $G(\pi)$ and $c_2(\pi)$ is the number of red edges in $2$-cycles of $G(\pi)$.
-Therefore, $b(\pi) - 2 c_2(\pi)$ is the number of \textit{free} red edges in $G(\pi)$ not occurring in $2$-cycles.
+Therefore, $b(\pi) - 2 c_2(\pi)$ is the number of red edges in $G(\pi)$ not occurring in $2$-cycles.
 $3 c_{3*}(\pi)$ is a lower bound to the number of red edges in $G(\pi)$ not occurring in $2$-cycles as every such cycle has at least three constituting red edges, proving the lemma.
@@
 
-\begin{example}
 Consider the alternating cycle decomposition
@@
 \end{align*}
-of the breakpoint graph of $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.
-\end{example}
+of the breakpoint graph of $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.\padding
 
@@
 
-\begin{example}
 Applying the reversal represented by vertex $\{4,5\}$ in figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
-\end{example}
 

### 2020-07-12 paper.tex
update abstract
update paper

@@
 The length of such a sequence is said to be the reversal distance of that permutation.
-In this paper we introduce fundamental concepts used in the literature on this topic.
+In this paper we introduce important concepts used in the literature on this topic.
 In particular, we introduce the fundamental relationship between the reversal distance and the number of alternating cycles in a cycle decomposition of a suitably-defined bi-colored graph.
 Then, a $3/2$-approximation is discussed in detail.
-We also reference further related research.
 \end{abstract}
@@
 A genome can be represented by a permutation.
-An inversion then simply is the reversal of a substring of a permutation \cite{Kececioglu1995}.
+An inversion then simply is the reversal of a substring of a permutation representing a genome \cite{Kececioglu1995}.
 
@@
 
-Perhaps the most natural algorithm for sorting by reversals is the \textit{ratchet algorithm} proposed by \citeauthor*{Watterson19821} \cite{Watterson19821}.
+Perhaps the most intuitive algorithm for sorting by reversals is the \textit{ratchet algorithm} proposed by \citeauthor*{Watterson19821} \cite{Watterson19821}.
 The algorithm sequentially brings all elements from $1$ to $n$ into the correct position.
@@
 
-We now try to find a tighter lower bound by introducing the breakpoint graph of a permutation first defined by \citeauthor*{Bafna1996}.
+We now try to find a tighter lower bound by introducing the breakpoint graph of a permutation that was first defined by \citeauthor*{Bafna1996}.
 
@@
 We connect two vertices $\pi_i$ and $\pi_j$ with a \textit{red} edge if they form a breakpoint in $\pi$.
-And we connect two vertices $\pi_i$ and $\pi_j$ representing non-consecutive elements in $\pi$ (that is $|\pi_i^{-1} - \pi_j^{-1}| \neq 1$) with a \textit{blue} edge if $\pi_i \sim \pi_j$.
+We connect two vertices $\pi_i$ and $\pi_j$ representing non-consecutive elements in $\pi$ (that is $|\pi_i^{-1} - \pi_j^{-1}| \neq 1$) with a \textit{blue} edge if $\pi_i \sim \pi_j$.
 \end{definition}
@@
 
-The red edges in the breakpoint graph identify the elements of $\pi$ that have to be split apart and the blue edges identify the elements of $\pi$ that have to be brought together in the process of sorting the permutation $\pi$ (the missing adjacencies).
+The red edges in the breakpoint graph represent the elements of $\pi$ that have to be split apart and the blue edges represent the elements of $\pi$ that have to be brought together in the process of sorting the permutation $\pi$ (the missing adjacencies).
 Figure \ref{fig:breakpoint_graph} shows the breakpoint graph of the permutation $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$.
@@
 An \textit{alternating cycle decomposition} of $G(\pi)$ is a set of edge-disjoint alternating cycles, such that every edge of $G(\pi)$ is contained in exactly one cycle of the set.
-We observe that for every breakpoint, a vertice in $G(\pi)$ must have a missing adjacency.
-And conversely, for every missing adjacency a vertex in $G(\pi)$ must form a breakpoint.
+We observe that for every breakpoint, a vertex in $G(\pi)$ must have a missing adjacency.
+Conversely, for every missing adjacency a vertex in $G(\pi)$ must form a breakpoint.
 Therefore, each vertex in our breakpoint graph has an equal number of incident red and blue edges.
@@
 
-The main result we prove in this section, relates the maximum number of alternating cycles of any cycle decomposition of $G(\pi)$ to the reversal distance of $\pi$.\padding
+The main result we prove in this section relates the maximum number of alternating cycles of any cycle decomposition of $G(\pi)$ to the reversal distance of $\pi$.\padding
 
@@
 An alternating cycle in $G(\pi)$ is a \textit{$k$-cycle} if it has $k$ constituting red and $k$ constituting blue edges.
-We say, a reversal \textit{acts on} two red edges of $G(\pi)$ if those two red edges represent the breakpoints that are split apart by the reversal.
-Now, an alternating cycle in $G(\pi)$ is \textit{oriented} if there is a $1$- or a $2$-reversal acting on two constituting red edges.
+We say a reversal \textit{acts on} two red edges of $G(\pi)$ if those two red edges represent the breakpoints that are split apart by the reversal.
+Now, an alternating cycle in $G(\pi)$ is \textit{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges.
 An alternating cycle of $G(\pi)$ is \textit{unoriented} if it is not oriented.
@@
 An alternating cycle of $G(\pi)$ is called \textit{undirected} if it is not directed.
-For example, in figure \ref{fig:breakpoint_graph} the alternating cycle $(\{0,7\},\{6,7\},\{3,6\},\{3,4\},\{1,4\},\{0,1\})$ is directed with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$, whereas the alternating cycle $(\{5,7\},\{4,5\},\{2,4\},\{1,2\},\{1,8\},\{7,8\})$ is undirected with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$.
+For example, in Figure \ref{fig:breakpoint_graph} the alternating cycle $(\{0,7\},\{6,7\},\{3,6\},\{3,4\},\{1,4\},\{0,1\})$ is directed with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$, whereas the alternating cycle $(\{5,7\},\{4,5\},\{2,4\},\{1,2\},\{1,8\},\{7,8\})$ is undirected with respect to $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$.
 \citeauthor*{Caprara1997} argues that an alternating cycle of $G(\pi)$ is oriented if and only if it is undirected \cite{Caprara1997}.
@@
 Let $i \sim i'$ and $j \sim j'$.
-We observe that the $2$-cycles shown in figure \ref{fig:2cycles} are in fact the only types of $2$-cycles (with the exception of the symmetric case of an oriented $2$-cycle).
+We observe that the $2$-cycles shown in Figure \ref{fig:2cycles} are in fact the only types of $2$-cycles (with the exception of the symmetric case of an oriented $2$-cycle).
 Given an oriented $2$-cycle, we know that the corresponding reversal acting on its two constituting red edges brings $j'$ next to $j$ and $i$ next to $i'$, removing two breakpoints.
@@
             \item Let $j \sim j'$.
-Then figure \ref{fig:case_i} illustrates that every alternating cycle in $G(\pi \circ \rho)$ containing the edge $\{i,k\}$ induces an alternating cycle in $G(\pi)$ containing the edges $\{j,i\}$, $\{j,j'\}$, and $\{j',k\}$.
+Then Figure \ref{fig:case_i} illustrates that every alternating cycle in $G(\pi \circ \rho)$ containing the edge $\{i,k\}$ induces an alternating cycle in $G(\pi)$ containing the edges $\{j,i\}$, $\{j,j'\}$, and $\{j',k\}$.
 Therefore, every alternating cycle decomposition of $G(\pi \circ \rho)$ into $c$ cycles induces an alternating cycle decomposition of $G(\pi)$ into $c$ cycles.
             \item Let $i \sim j$, $i \sim k$ and $j \sim l$.
-Then figure \ref{fig:case_ii} illustrates that every alternating cycle in $G(\pi \circ \rho)$ containing the edge $\{i,j\}$ induces an alternating cycle in $G(\pi)$ containing the edges $\{i,k\}$, $\{k,l\}$, and $\{j,l\}$.
+Then Figure \ref{fig:case_ii} illustrates that every alternating cycle in $G(\pi \circ \rho)$ containing the edge $\{i,j\}$ induces an alternating cycle in $G(\pi)$ containing the edges $\{i,k\}$, $\{k,l\}$, and $\{j,l\}$.
 Therefore, every alternating cycle decomposition of $G(\pi \circ \rho)$ into $c$ cycles induces an alternating cycle decomposition of $G(\pi)$ into $c$ cycles.
@@
 
-Using this lower bound, \citeauthor*{Bafna1996} already describe a $7/4$-approximation that is not considered in more detail here.
+Using this lower bound, \citeauthor*{Bafna1996} already describe a $7/4$-approximation that is not considered here.
 \citeauthor*{Christie1998} formulated a lower bound in terms of the number of $2$-cycles in a maximum alternating cycle decomposition \cite{Christie1998}.
@@
 In the case of a $2$-cycle, two vertices in the reversal graph represent the same reversal.
-A vertex is colored red if the represented reversal is a $1$- or a $2$-reversal.
+A vertex is colored red if the represented reversal is a $1$- or $2$-reversal.
 Otherwise a vertex is colored blue.
@@
 \end{align*}
-of the breakpoint graph of $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$. The corresponding reversal graph is given in figure \ref{fig:reversal_graph}.\padding
+of the breakpoint graph of $(0\ 7\ 5\ 6\ 3\ 2\ 4\ 1\ 8)$. The corresponding reversal graph is given in Figure \ref{fig:reversal_graph}.\padding
 
 We observe that the only alternating cycle decomposition of $G(id)$ is $\mathcal{C} = \emptyset$.
-And therefore, the only reversal graph of the identity permutation $R(\emptyset)$ consists of $n+1$ isolated blue vertices.
+Therefore, the only reversal graph of the identity permutation $R(\emptyset)$ consists of $n+1$ isolated blue vertices.
 In the remainder of this section, our goal is to describe a way of transforming the reversal graph of a given alternating cycle decomposition into the reversal graph only consisting of isolated blue vertices with as few reversals as possible.
@@
 
-Applying the reversal represented by vertex $\{4,5\}$ in figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
+Applying the reversal represented by vertex $\{4,5\}$ in Figure \ref{fig:reversal_graph} results in the breakpoint and reversal graphs shown in Figures \ref{fig:breakpoint_graph_change} and \ref{fig:reversal_graph_change} respectively.
 
@@
 We call a $2$-cycle of $G(\pi)$ \textit{selected} if its corresponding edge of $F(\pi)$ is part of the matching $M$.
-Note, that we haven't yet found the set of edge-disjoint $2$-cycles that we need to construct an alternating cycle decomposition.
+Note, that we have not yet found the set of edge-disjoint $2$-cycles that we need to construct an alternating cycle decomposition.
 This is because the matching ensures that all selected $2$-cycles are edge-disjoint on their red edges, but does not ensure that they are also edge-disjoint on their blue edges.
@@
 Let $C$ be an unoriented ladder $2$-cycle in $\mathcal{C}$.
-Then the vertices of $R(\mathcal{C})$ that represent $C$ are part of a component that contains vertices representing an alternating cycle that is not a selected $2$-cycle.
+Then the vertices of $R(\mathcal{C})$ that represent $C$ are part of a connected component that contains vertices representing an alternating cycle that is not a selected $2$-cycle.
 \end{lemma}
@@
 
-The approximation algorithm with a performance ratio of $\frac{3}{2}$, is outlined in figure \ref{alg:approximation}.
+The approximation algorithm with a performance ratio of $\frac{3}{2}$, is outlined in Figure \ref{alg:approximation}.
 
@@
 The reversal graph can be constructed in $O(n^2)$ time.
-The algorithm to find an elimination sequence of $R(\mathcal{C})$ is outlined in figure \ref{alg:elimination_sequence} and runs in $O(n^4)$ time.
+The algorithm to find an elimination sequence of $R(\mathcal{C})$ is outlined in Figure \ref{alg:elimination_sequence} and runs in $O(n^4)$ time.
 So the overall time-complexity of the algorithm is $O(n^4)$ \cite{Christie1998}.

### 2020-07-13 paper.tex
update paper

@@
 \begin{abstract}
-    A reversal is a transformation, reversing the order of elements of a permutation.
+    A reversal is a transformation that inverses the order of elements of a permutation.
 Sorting by reversals is the problem of finding the shortest sequence of reversals transforming a permutation to the identity.
@@
 \begin{align*}
-    (1\ \cdots\ i-1\ j\ j-1\ \cdots\ i+1\ i\ j+1\ \cdots\ n).
+    (1\ \cdots\ i-1\ \ j\ \ j-1\ \cdots\ i+1\ \ i\ \ j+1\ \cdots\ n).
 \end{align*}
@@
 
-We now consider a couple of lemmas that will help use prove the approximation bound later.
+We now consider a couple of lemmas that will later help us in proving the approximation bound.
 

### 2020-07-13 paper.tex
spelling

@@
 In this paper we give an overview of approximation algorithms for MIN-SBR.
-We first introduce the notion of breakpoints given by \citeauthor*{Kececioglu1995} allowing us to formulate a first lower bound for the reversal distance of $\pi$ and leading to a $2$-approximation \cite{Kececioglu1995} (section \ref{sec:breakpoints}).
-In section \ref{sec:breakpoint_graph} we define a suitable bi-colored graph first introduced by \citeauthor*{Bafna1996} to find a tighter lower bound \cite{Bafna1996}.
+We first introduce the notion of breakpoints given by \citeauthor*{Kececioglu1995} allowing us to formulate a first lower bound for the reversal distance of $\pi$ and leading to a $2$-approximation \cite{Kececioglu1995} (Section \ref{sec:breakpoints}).
+In Section \ref{sec:breakpoint_graph} we define a suitable bi-colored graph first introduced by \citeauthor*{Bafna1996} to find a tighter lower bound \cite{Bafna1996}.
 In doing so, we make a fundamental observation relating the reversal distance to the number of alternating cycles in a cycle decomposition of such a graph.
@@
 
-First, we need to define some terminology.
+First, we introduce some terminology.
 An alternating cycle in $G(\pi)$ is a \textit{$k$-cycle} if it has $k$ constituting red and $k$ constituting blue edges.
@@
 
-A similar lemma is proven by \citeauthor*{Christie1998} which we use as a corollary of this lemma in section \ref{sec:reversal_graph} \cite{Christie1998}.\padding
+A similar lemma is proven by \citeauthor*{Christie1998} which we use as a corollary of this lemma in Section \ref{sec:reversal_graph} \cite{Christie1998}.\padding
 
@@
 
-Using \ref{thm:1}, \citeauthor*{Bafna1996} give a lower bound of the reversal distance of a permutation $\pi$ in terms of $b(\pi)$ and $c(\pi)$ \cite{Bafna1996}.
+Using Theorem \ref{thm:1}, \citeauthor*{Bafna1996} give a lower bound of the reversal distance of a permutation $\pi$ in terms of $b(\pi)$ and $c(\pi)$ \cite{Bafna1996}.
 
@@
 \begin{proof}
-Follows immediately from theorem \ref{thm:2} and lemma \ref{lem:2}.
+Follows immediately from Theorem \ref{thm:2} and Lemma \ref{lem:2}.
 \end{proof}
@@
 In this section we discuss the $3/2$-approximation proposed by \citeauthor*{Christie1998} in detail \cite{Christie1998}.
-First, we describe how a sorting sequence of reversals can be obtained from an alternating cycle decomposition of $G(\pi)$ (section \ref{sec:reversal_graph}).
-In section \ref{sec:cycle_decomposition} we then describe how to find such a cycle decomposition.
-This then allows us to prove the approximation bound (section \ref{sec:approximation_bound}).
-Lastly, we describe the full algorithm and analyze its complexity (section \ref{sec:algorithm}).
+First, we describe how a sorting sequence of reversals can be obtained from an alternating cycle decomposition of $G(\pi)$ (Section \ref{sec:reversal_graph}).
+In Section \ref{sec:cycle_decomposition} we then describe how to find such a cycle decomposition.
+This then allows us to prove the approximation bound (Section \ref{sec:approximation_bound}).
+Lastly, we describe the full algorithm and analyze its complexity (Section \ref{sec:algorithm}).
 
@@
 \begin{proof}
-As $A$ is oriented, we can use lemma \ref{lem:7} to repeatedly find a red vertex $u$ in $A$ such that $A_u$ is still oriented and apply $\rho(u)$.
-By lemma \ref{lem:6}, we know that every vertex in $R(\mathcal{C})$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component.
+As $A$ is oriented, we can use Lemma \ref{lem:7} to repeatedly find a red vertex $u$ in $A$ such that $A_u$ is still oriented and apply $\rho(u)$.
+By Lemma \ref{lem:6}, we know that every vertex in $R(\mathcal{C})$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component.
 Therefore, every alternating cycle represented by vertices in $A$ eventually reduces to a $2$-cycle.
-Moreover, by lemma \ref{lem:1}, $\rho(u)$ is a $2$-reversal iff $u$ arises from an oriented $2$-cycle.
+Moreover, by Lemma \ref{lem:1}, $\rho(u)$ is a $2$-reversal iff $u$ arises from an oriented $2$-cycle.
 \end{proof}
@@
 \begin{proof}
-Let $u$ be any blue vertex of $A$, then $\rho(u)$ is a $0$-reversal. By lemma \ref{lem:3}, $A_u$ is oriented.
-Now, apply lemma \ref{lem:8} to obtain an elimination sequence for $A_u$.
+Let $u$ be any blue vertex of $A$, then $\rho(u)$ is a $0$-reversal. By Lemma \ref{lem:3}, $A_u$ is oriented.
+Now, apply Lemma \ref{lem:8} to obtain an elimination sequence for $A_u$.
 \end{proof}
@@
 
-By theorem \ref{thm:3}, an algorithm that finds a sorting sequence of reversals of at most length $b(\pi) - \frac{1}{2} c_2(\pi)$ achieves an approximation bound of $\frac{3}{2}$.
+By Theorem \ref{thm:3}, an algorithm that finds a sorting sequence of reversals of at most length $b(\pi) - \frac{1}{2} c_2(\pi)$ achieves an approximation bound of $\frac{3}{2}$.
 
@@
 \begin{proof}
-Using theorem \ref{thm:5}, first find an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$ with at least $\left\lceil \frac{y}{2} \right\rceil$ ladder $2$-cycles and $z$ independent $2$-cycles.
+Using Theorem \ref{thm:5}, first find an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$ with at least $\left\lceil \frac{y}{2} \right\rceil$ ladder $2$-cycles and $z$ independent $2$-cycles.
 
 Let $k$ be the number of $2$-cycles in oriented connected components of $R(\mathcal{C})$.
-Let $u$ be the number of unoriented connected components in $R(\mathcal{C})$ that include $l$ selected $2$-cycles and that, by lemma \ref{lem:10}, contain vertices representing remaining unselected $2$-cycles.
+Let $u$ be the number of unoriented connected components in $R(\mathcal{C})$ that include $l$ selected $2$-cycles and that, by Lemma \ref{lem:10}, contain vertices representing remaining unselected $2$-cycles.
 Let $v$ be the number of remaining unoriented connected components consisting only of vertices representing $m$ independent selected $2$-cycles.
@@
 
-By theorem \ref{thm:4}, we can now sort $\pi$ using at least $k+l+u+m$ $2$-reversals and only $u+v$ $0$-reversals. Therefore
+By Theorem \ref{thm:4}, we can now sort $\pi$ using at least $k+l+u+m$ $2$-reversals and only $u+v$ $0$-reversals. Therefore
 \begin{align*}
@@
 As $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles in $\mathcal{C}$, $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ holds.
-    \item As each connected components of $R(\mathcal{C})$ counted by $v$ represents at least one unoriented $2$-cycle, lemma \ref{lem:5} implies that such a connected component must represent at least two alternating cycles in $\mathcal{C}$.
+    \item As each connected components of $R(\mathcal{C})$ counted by $v$ represents at least one unoriented $2$-cycle, Lemma \ref{lem:5} implies that such a connected component must represent at least two alternating cycles in $\mathcal{C}$.
 As the connected component only represents independent $2$-cycles, we know that it must represent at least two independent $2$-cycles.

### 2020-07-14 paper.tex
update paper abstract

@@
 \begin{abstract}
-    A reversal is a transformation that inverses the order of elements of a permutation.
+    A reversal is a transformation that inverses the order of some consecutive elements of a permutation.
 Sorting by reversals is the problem of finding the shortest sequence of reversals transforming a permutation to the identity.
@@
 In particular, we introduce the fundamental relationship between the reversal distance and the number of alternating cycles in a cycle decomposition of a suitably-defined bi-colored graph.
-Then, a $3/2$-approximation is discussed in detail.
+Then, a $3/2$-approximation obtained by \citeauthor*{Christie1998} is discussed in detail.
 \end{abstract}
@@
 The reversal distance is denoted by $d(\pi)$.
-We say that a sequence of reversals transforming $\pi$ into $id$, \textit{sorts} $\pi$.
+A sequence of reversals transforming $\pi$ into $id$, \textit{sorts} $\pi$.
 \end{definition}

### 2021-03-26 slides.tex
paper
update slides date
deploy: 1583efd39cc6fce7077825fe6a927d75ff028beb

@@
 \author{Jonas Hübotter}
+\date{July 10, 2020}
 
@@
 \label{thm:3}
-Let $\pi \in S_n$ and $C$ be a maximum alternating cycle decomposition of $G(\pi)$. Let $\g{c_2(\pi)}$ be the minimum number of alternating $2$-cycles in any such $C$. Then
+Let $\pi \in S_n$ and $\mathcal{C}$ be a maximum alternating cycle decomposition of $G(\pi)$. Let $\g{c_2(\pi)}$ be the minimum number of alternating $2$-cycles in any such $\mathcal{C}$. Then
 \begin{align*}
@@
 \begin{enumerate}
-    \item given an alternating cycle decomposition $C$ of $G(\pi)$ we find a sorting sequence of reversals for $\pi$\pause; and
+    \item given an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$ we find a sorting sequence of reversals for $\pi$\pause; and
     \item we find an alternating cycle decomposition of $G(\pi)$ maximizing the number of $2$-cycles.
@@
 \begin{definition}[reversal graph, \citeauthor*{Christie1998}]
-Given an alternating cycle decomposition $C$ of $G(\pi)$, \par\pause
-define $R(C)$ with
+Given an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$, \par\pause
+define $R(\mathcal{C})$ with
 \begin{itemize}
     \item an isolated \b{blue} vertex for each adjacency in $\pi$\pause;
-    \item $m$ vertices for each $m$-cycle in $C$ each representing the reversal $\rho(u)$ acting on the two red edges connected by a blue edge\pause;
+    \item $m$ vertices for each $m$-cycle in $\mathcal{C}$ each representing the reversal $\rho(u)$ acting on the two red edges connected by a blue edge\pause;
     \begin{itemize}
@@
 
-Observation: The only alternating cycle decomposition of $G(id)$ is $C = \emptyset$.\pause
+Observation: The only alternating cycle decomposition of $G(id)$ is $\mathcal{C} = \emptyset$.\pause
 
@@
 Let $\pi = (0\ 1\ 3\ 4\ 2\ 5) \in S_4$. \par
-Given the alternating cycle decomposition $C$ of $G(\pi)$
+Given the alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$
 
@@
 {\footnotesize\begin{align*}
-    C = \{(&\{1,3\},\{2,3\},\{2,4\}, \\
+    \mathcal{C} = \{(&\{1,3\},\{2,3\},\{2,4\}, \\
                      &\{4,5\},\{2,5\},\{1,2\})\}
@@
 
-construct $R(C)$.
+construct $R(\mathcal{C})$.
 
@@
 
-Idea: Each connected component of $R(C)$ can be sorted separately.
-
-\end{frame}
-
-\begin{frame}
-
-Let $u$ be a vertex of $R(C)$ representing reversal $\rho(u)$.\pause
+Idea: Each connected component of $R(\mathcal{C})$ can be sorted separately.
+
+\end{frame}
+
+\begin{frame}
+
+Let $u$ be a vertex of $R(\mathcal{C})$ representing reversal $\rho(u)$.\pause
 
 \begin{definition}
-Denote by $C_u$ the alternating cycle decomposition of $G(\pi \circ \rho(u))$ that is obtained from $C$.% by removing one 2-cycle or shortening one of its cycles by one red edge.
+Denote by $\mathcal{C}_u$ the alternating cycle decomposition of $G(\pi \circ \rho(u))$ that is obtained from $\mathcal{C}$.% by removing one 2-cycle or shortening one of its cycles by one red edge.
 \end{definition}\pause
@@
 \label{lem:5}
-$R(C_u)$ can be derived from $R(C)$ by making the following changes to $R(C)$:\pause
+$R(\mathcal{C}_u)$ can be derived from $R(\mathcal{C})$ by making the following changes to $R(\mathcal{C})$:\pause
 \begin{enumerate}
@@
 \begin{tabularx}{\textwidth}{X|R}
-$G(\pi)$ and $R(C)$ & $G(\pi \circ \rho(u))$ and $R(C_u)$ \\ \hline
+$G(\pi)$ and $R(\mathcal{C})$ & $G(\pi \circ \rho(u))$ and $R(\mathcal{C}_u)$ \\ \hline
 \spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
@@
 % \begin{corollary}[\citeauthor*{Christie1998}]
-% $R(C)$ contains no isolated blue vertices arising from cycles in $C$.
+% $R(\mathcal{C})$ contains no isolated blue vertices arising from cycles in $\mathcal{C}$.
 % \end{corollary}
@@
 \label{lem:3}
-All vertices arising from the same alternating cycle in $C$ are in the same connected component of $R(C)$.
+All vertices arising from the same alternating cycle in $\mathcal{C}$ are in the same connected component of $R(\mathcal{C})$.
 \end{lemma}\pause
@@
 \label{lem:7}
-Vertices arising from an unoriented $2$-cycle of $C$ must be in a connected component of $R(C)$ with vertices arising from at least one more alternating cycle of $C$.
+Vertices arising from an unoriented $2$-cycle of $\mathcal{C}$ must be in a connected component of $R(\mathcal{C})$ with vertices arising from at least one more alternating cycle of $\mathcal{C}$.
 \end{lemma}
@@
 % \begin{proof}
-% Let $u$ and $v$ represent the blue edges of an unoriented $2$-cycle of $C$. \par\pause
-% Then its two constituting blue edges do not interleave. Hence, $R(C)$ does not connect $u$ and $v$ with an edge. \par\pause
-% If its constituting blue edges do not interleave with blue edges of any other alternating cycle of $C$, then $u$ and $v$ would be isolated blue vertices. \b{$\lightning$}
+% Let $u$ and $v$ represent the blue edges of an unoriented $2$-cycle of $\mathcal{C}$. \par\pause
+% Then its two constituting blue edges do not interleave. Hence, $R(\mathcal{C})$ does not connect $u$ and $v$ with an edge. \par\pause
+% If its constituting blue edges do not interleave with blue edges of any other alternating cycle of $\mathcal{C}$, then $u$ and $v$ would be isolated blue vertices. \b{$\lightning$}
 % \end{proof}
@@
 \begin{definition}
-We call a connected component of $R(C)$ \b{oriented} if it contains a red vertex or if it consists solely of an isolated blue vertex. \pause\spadding
-
-Let $A$ be a connected component of $R(C)$. We denote by $A_u$ the subgraph of $R(C_u)$ that contains all the vertices of $A$.
+We call a connected component of $R(\mathcal{C})$ \b{oriented} if it contains a red vertex or if it consists solely of an isolated blue vertex. \pause\spadding
+
+Let $A$ be a connected component of $R(\mathcal{C})$. We denote by $A_u$ the subgraph of $R(\mathcal{C}_u)$ that contains all the vertices of $A$.
 \end{definition}\pause
@@
 \label{lem:4}
-If a connected component $A$ of $R(C)$ is oriented and not an isolated blue vertex, it contains a red vertex $u$ such that $A_u$ is still oriented.
+If a connected component $A$ of $R(\mathcal{C})$ is oriented and not an isolated blue vertex, it contains a red vertex $u$ such that $A_u$ is still oriented.
 \end{lemma}
@@
 % \begin{definition}
-% A sequence of reversals that turns the connected component $A$ of $R(C)$ into an isolated blue vertex, is called an \b{elemination sequence} for $A$.
+% A sequence of reversals that turns the connected component $A$ of $R(\mathcal{C})$ into an isolated blue vertex, is called an \b{elemination sequence} for $A$.
 % \end{definition}\pause
@@
 % \label{lem:6}
-% Let $A$ be an oriented connected component of $R(C)$ arising from $k$ different alternating cycles of $G(\pi)$.
+% Let $A$ be an oriented connected component of $R(\mathcal{C})$ arising from $k$ different alternating cycles of $G(\pi)$.
 Then there is an \textit{elimination sequence} for $A$ that contains $k$ $2$-reversals with all the other reversals being $1$-reversals.
@@
 % \begin{itemize}
-%     \item every vertex in $R(C)$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component (lemma \ref{lem:3})\pause;
+%     \item every vertex in $R(\mathcal{C})$ arising from an alternating cycle of $G(\pi)$ is contained within the same connected component (lemma \ref{lem:3})\pause;
 %     \item  as $A$ is oriented, we can use lemma \ref{lem:4} to repeatedly find a red vertex $u$ in $A$ and apply $\rho(u)$\pause; and
@@
 % \begin{lemma}[\citeauthor*{Christie1998}]
-% Let $A$ be an unoriented connected component of $R(C)$ arising from $k$ different alternating cycles of $G(\pi)$.
+% Let $A$ be an unoriented connected component of $R(\mathcal{C})$ arising from $k$ different alternating cycles of $G(\pi)$.
 Then there is an \textit{elimination sequence} for $A$ that contains one $0$-reversal, $k$ $2$-reversals, with all the other reversals being $1$-reversals.
@@
 \label{thm:4}
-Let $\pi \in S_n$ and $C$ be a cycle decomposition of $G(\pi)$. Then
-\begin{align*}
-    d(\pi) \leq \r{b(\pi)} - |C| + u(C)
+Let $\pi \in S_n$ and $\mathcal{C}$ be an alternating cycle decomposition of $G(\pi)$. Then
+\begin{align*}
+    d(\pi) \leq \r{b(\pi)} - |\mathcal{C}| + u(\mathcal{C})
 \end{align*}
-where $u(C)$ is the number of unoriented components in $R(C)$.
+where $u(\mathcal{C})$ is the number of unoriented components in $R(\mathcal{C})$.
 \end{theorem}
@@
 \label{thm:5}
-Given a maximum cardinality matching $M$ of $F(\pi)$ it is possible to find a cycle decomposition $C$ of $G(\pi)$ that contains at least $\left\lceil \frac{y}{2} \right\rceil$ ladder $2$-cycles and $z$ independent $2$-cycles.
+Given a maximum cardinality matching $M$ of $F(\pi)$ it is possible to find an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$ that contains at least $\left\lceil \frac{y}{2} \right\rceil$ ladder $2$-cycles and $z$ independent $2$-cycles.
 \end{theorem}
@@
 % \begin{proof}
-% Let $C$ contain all independent $2$-cycles from $L(M)$ and every alternate cycle of each ladder. \par\pause
-% Obtain the rest of $C$ by adding any alternating cycle decomposition of the remaining unused edges of $G(\pi)$.
+% Let $\mathcal{C}$ contain all independent $2$-cycles from $L(M)$ and every alternate cycle of each ladder. \par\pause
+% Obtain the rest of $\mathcal{C}$ by adding any alternating cycle decomposition of the remaining unused edges of $G(\pi)$.
 % \end{proof}
@@
 \begin{proof}\renewcommand{\qedsymbol}{}
-Using theorem \ref{thm:5}, first find an alternating cycle decomposition $C$ of $G(\pi)$ with at least $\left\lceil \frac{y}{2} \right\rceil$ $2$-cycles as part of ladders and $z$ independent $2$-cycles.
+Using theorem \ref{thm:5}, first find an alternating cycle decomposition $\mathcal{C}$ of $G(\pi)$ with at least $\left\lceil \frac{y}{2} \right\rceil$ $2$-cycles as part of ladders and $z$ independent $2$-cycles.
 \end{proof}
@@
 \begin{itemize}
-    \item Let $k$ be the number of $2$-cycles in oriented connected components of $R(C)$.\pause
-    \item Let $u$ be the number of unoriented connected components in $R(C)$ that include $l$ selected $2$-cycles and that contain vertices representing remaining unselected $2$-cycles.\pause
+    \item Let $k$ be the number of $2$-cycles in oriented connected components of $R(\mathcal{C})$.\pause
+    \item Let $u$ be the number of unoriented connected components in $R(\mathcal{C})$ that include $l$ selected $2$-cycles and that contain vertices representing remaining unselected $2$-cycles.\pause
     \item Let $v$ be the number of remaining unoriented connected components consisting only of vertices representing $m$ independent selected $2$-cycles.
@@
 \begin{enumerate}
-    \item $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ as $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles in $C$\pause;
+    \item $k+l+m \geq \left\lceil \frac{y}{2} \right\rceil + z$ as $\left\lceil \frac{y}{2} \right\rceil + z$ is the number of selected $2$-cycles in $\mathcal{C}$\pause;
     \item $v \leq \left\lfloor \frac{z}{2} \right\rfloor$ as every unoriented component representing a $2$-cycle represents at least one more alternating cycle (lemma \ref{lem:7})\pause; and
@@
 
-\nocite{*}
 \printbibliography

### 2021-03-26 min-sbr.hs

@@ -1,0 +1,217 @@
+module MIN_SBR where
+
+
+import Data.List ((\\),elemIndex)
+import Data.Maybe (fromJust)
+import Test.QuickCheck
+
+import Data.Foldable (find)
+import Data.Maybe (isJust)
+import Control.Monad (join)
+
+firstJust = join . find isJust
+
+
+type Element = Int
+newtype Permutation = Permutation [Element] deriving (Show, Eq)
+
+-- Reversal leftRedEdge rightRedEdge blueEdge
+data Reversal = Reversal (Element,Element) (Element,Element) (Element,Element) deriving (Show, Eq)
+
+-- BreakpointGraph vertices redEdges blueEdges
+type BreakpointGraphEdge = (Element,Element)
+data BreakpointGraph = BreakpointGraph [Element] [BreakpointGraphEdge] [BreakpointGraphEdge] deriving (Show, Eq)
+data CycleDecomposition = CycleDecomposition [[BreakpointGraphEdge]] deriving (Show, Eq)
+
+-- ReversalGraph vertices edges
+type ReversalGraphEdge = (BreakpointGraphEdge,BreakpointGraphEdge)
+data ReversalGraph = ReversalGraph [BreakpointGraphEdge] [ReversalGraphEdge] deriving (Show, Eq)
+
+sorted :: (Element,Element) -> (Element,Element)
+sorted (a,b)
+  | a <= b    = (a,b)
+  | otherwise = (b,a)
+
+isId :: Permutation -> Bool
+isId (Permutation xs) = xs == [0..(length xs) - 1]
+
+isIdR :: ReversalGraph -> Bool
+isIdR (ReversalGraph _ xs) = null xs
+
+isAdjacency :: Element -> Element -> Bool
+isAdjacency x y = abs (x - y) == 1
+
+isBreakpoint :: Element -> Element -> Bool
+isBreakpoint x y = not (isAdjacency x y)
+
+neighbors :: [Element] -> [(Element,Element)]
+neighbors xs = map (\x -> (x,x+1)) [0..(length xs) - 2]
+
+breakpoints :: [Element] -> [(Element,Element)]
+breakpoints [x] = []
+breakpoints (x:y:zs)
+  | isBreakpoint x y = sorted (x,y) : breakpoints (y:zs)
+  | otherwise        = breakpoints (y:zs)
+
+adjacencies :: [Element] -> [(Element,Element)]
+adjacencies [x] = []
[…]
